### 2024-10-20 paper/paper.qmd
Update paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Robert Ford
  - Michelle Ji
  - Cher Ning
thanks: "Code and data are available at: https://github.com/Ford-Robert/us-presidential-election."
date: today
date-format: long
abstract: "Compiled polling surveys regarding the 2024 United States presidential election were analyzed in this report based on data from FiveThirtyEight [@538data]. From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system. The models forecast Kamala Harris as the winner of the US presidency. Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process. Additionally, it helps predict potential policy shifts and holds political institutions accountable, contributing to a more transparent and informed democracy."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(kableExtra)
library(ggplot2)
library(here)
library(kableExtra)

#read in cleaned data
cleaned_poll_data <-
  read_csv(file = here("data/cleaned_poll_data.csv"), 
           show_col_types = FALSE)
```

# Introduction {#sec-intro}

Every four years, the United States presidential election occurs and is one of the most significant political events internationally. The complex process of the election involves both the popular and electoral college system, in which voters cast ballots on election day in each state. The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege]. After the popular vote is accounted for, the candidate who receives the majority of the electoral votes from all states wins the presidential election. The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege]. Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs. For example, regarding technology, both parties are concerned with the rapid progression of artificial intelligence’s capabilities but differ in the response and regulation of it. Republicans believe that the current practices are unfair and disproportionately target their voices, whereas democrats generally are in favor of tighter regulation [@keyissues]. Another key issue important to voters and the candidates is abortion. As former president, he supported restrictions on abortion access and advocated for policies that align with the conservative movement’s opposition to abortion rights [@abortionstance]. On the other hand, Harris has been a key supporter in the pro-choice movement and has been an advocate for protecting and expanding access to abortion services, as well as female autonomy [@harrisviews]. The outcome of this election will have profound consequences for various legislative policies and the future direction of the United States.

To forecast the 2024 United States presidential election, a generalized linear model and bayesian model were built. Our goal is to determine who will win the election based on the electoral college system. A key finding from our model is that we predict Kamala Harris to win the presidential election. This would be a historic win, as she would be the first woman to become US president. This would represent a significant milestone in breaking barriers related to gender and racial representation in the highest levels of American politics, symbolizing progress toward a more inclusive and diverse political landscape. One drawback of our model stems from representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.

This paper is broken down into various sections, including Data, Modeling, Results, Discussion, Conclusion, and Appendices. This paper uses data made available by FiveThirtyEight [@538data] which compiles individual polling surveys based on state and methodology. @sec-data explores the data, highlighting key aspects that may be useful to future policymakers or campaign strategists, as well as details the variables present in the dataset. @sec-model presents the models that were built and used to forecast the election. @sec-results details the conclusions of our model and @sec-discussion explores possible implications and insights of our conclusions. @sec-appendices include an idealized methodology and survey that we could hypothetically run, with the task of forecasting the US presidential election. This section also includes an in-depth analysis of one specific pollster’s methodology found within the larger dataset.

# Data {#sec-data}
The dataset used was obtained through FiveThirtyEight Interactives [@538data] and is focused on the United States presidential general polls for 2024. The polling data extracted are polling averages, where FiveThirtyEight [@538] includes individual polling data from all publicly available polls that meet their ethical and methodological standards and at minimum, must test Harris versus Trump, the two primary presidential nominees for each political party. From the individual polls, FiveThirtyEight [@538] will reweight and adjust the polls based on a few criteria, such as poll sample size, how recent it is, and house effects. The individual poll data is then added to the dataset, which is accessible for us to use. This dataset is crucial to help us build a model to forecast the 2024 US Presidential election and can help us analyze key features in voter habits for future political analysis.

There were two other potential datasets that could have been used, titled “Presidential Polling Averages” and “President Primary Polls”, both found through FiveThirtyEight [@538]. The “Presidential Polling Averages” dataset only included data from the Biden versus Trump election in 2020 and was not updated to include data for the upcoming 2024 election and therefore inadequate. The “Presidential Primary Polls” dataset included very similar information as our chosen dataset, however, it included candidates that have since dropped out or were not major running candidates, such as Nikki Haley and Michelle Obama. It includes observations that are unnecessary and creates noise, therefore we did not choose this dataset either.

The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region. Pollster Name indicates which pollster the individual poll came from, which varies from individual companies, such as YouGov, to research institutions, such as Quinnipiac University and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time. This is to adjust for any bias. Method is the way the poll was deployed, including but not limited to live phone, online panel, and text-to-web. State indicates which state the poll was conducted. State indicates what state the poll was conducted and sample size is the size of the poll. The pollscore is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number. The election date and days to election variables indicate how far out the poll was conducted with respect with the election date of November 5, 2024. We also reorganized the dataset to include a poll region, which divides the US into geographical regions such as rust belt, northeast, etc.

@table-1 details the first few observations of the cleaned dataset, which we used to build our model.

```{r}
#| label: table-1
#| tbl-cap: First few observations of cleaned data set
#| echo: false

head(cleaned_poll_data, 5) |>
  kable(
    col.names = c("pollster", "numeric_grade", "pollscore",	"end_date",	"transparency_score", "question_id", "method", "state", "sample_size", "candidate", "support", "election_date", "days_to_election", "initial_weight", "weight", "pol_region"),
  booktabs = TRUE
  )
```

For further analysis, the cleaned dataset was separated into two different datasets by beach name, for simplicity purposes. Data was analyzed through the R programming software [@citeR] and packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], and dplyr [@dplyr] were used to help download, clean, simulate, analyze, and test the data.

Some pollsters are included in the dataset more than others based on the frequency of their conducted poll surveys and if they align with FiveThirtyEight’s guidelines, which can be seen in @fig-1. @fig-1 highlights the distribution of the top 25 pollsters and how frequently they appear in the dataset, with Morning Consult being the most popular pollster with almost 500 appearances. Noting which pollsters are more trustworthy and provide reliable data is important for evaluating the quality of future analyses.

```{r}
#| label: fig-1
#| fig-cap: Top 25 pollsters in overall polling data
#| echo: false
#| warning: false
#| message: false

pollster_counts <- cleaned_poll_data |>
  count(pollster, name = "n") %>%
  arrange(desc(n))

top_25_pollsters <- pollster_counts |>
  slice_head(n = 25)

ggplot(top_25_pollsters, aes(x = reorder(`pollster`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Pollster", y = "Number of Appearances", title = "Top 25 Pollsters") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6))
```

 @fig-2 details the frequency of various polling methods used in the individual polling data, showing online panels are the most popular methodology used to deploy these polls, followed by live phone. The polling method plays a crucial role in reducing certain response biases. For instance, if the poll is conducted via an online panel with anonymous responses, participants may feel more comfortable expressing their true voting intentions, potentially reducing social desirability bias.

```{r}
#| label: fig-2
#| fig-cap: Frequency of Various Polling Methods
#| echo: false
#| warning: false
#| message: false

polling_method_counts <- cleaned_poll_data |>
  count(method, name = "n") |>
  arrange(desc(n))

top_25_methods <- polling_method_counts |>
  slice_head(n = 25)

ggplot(top_25_methods, aes(x = reorder(`method`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(x = "Polling Method", y = "Number of Appearances", title = "Frequency of Polling Methods") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7))
```

Finally, @fig-3 shows the number of votes per candidate. As expected, Harris and Trump have the highest number of votes, with Harris having slightly more votes based on the surveyed polls than Trump. Other candidates with a noticeable number of votes in the graph include Robert F. Kennedy and Jill Stein. However, their totals are significantly lower than those of Harris and Trump, making them likely insignificant in the election. It is also important to mention that Biden still appears in these polls, and therefore in @fig-3, due to the lag time in poll data processing by FiveThirtyEight. However, since he has dropped out of the race, he will not be included in our model.

```{r}
#| label: fig-3
#| fig-cap: Number of Votes per Candidate
#| echo: false
#| warning: false
#| message: false

cleaned_poll_data <- cleaned_poll_data |> 
  mutate(sample_size = ifelse(is.na(sample_size), 0, sample_size))

votes = cleaned_poll_data$support / 100 * cleaned_poll_data$sample_size

poll_summary <- cleaned_poll_data |>
  group_by(candidate) |>
  summarize(votes = sum(votes, na.rm = TRUE))
poll_summary <- poll_summary |>
  filter(!is.na(candidate), !is.na(votes))

ggplot(cleaned_poll_data, aes(x = candidate, y = votes, fill = candidate)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Votes per Candidate", 
       x = "Candidate", 
       y = "Number of Votes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10))
```

# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {#sec-appendices}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-20 paper/paper.qmd
Update paper.qmd

@@
 format: pdf
+header-includes:
+  - \usepackage{geometry}
+  - \geometry{margin=0.75in} 
 number-sections: true
@@
 
-@table-1 details the first few observations of the cleaned dataset, which we used to build our model.
+The table details the first few observations of the cleaned dataset, which we used to build our model.
 
 ```{r}
-#| label: table-1
+#| label: tbl-1
 #| tbl-cap: First few observations of cleaned data set
@@
 
-head(cleaned_poll_data, 5) |>
-  kable(
-    col.names = c("pollster", "numeric_grade", "pollscore",	"end_date",	"transparency_score", "question_id", "method", "state", "sample_size", "candidate", "support", "election_date", "days_to_election", "initial_weight", "weight", "pol_region"),
-  booktabs = TRUE
-  )
+first_table <- cleaned_poll_data[, c("pollster", "numeric_grade", "pollscore", 
+                                       "end_date", "transparency_score", 
+                                       "question_id", "method")]
+
+kable(head(first_table, 5), col.names = names(first_table), 
+      booktabs = TRUE)
+
+second_table <- cleaned_poll_data[, c("state", "sample_size", "candidate", 
+                                        "support", "election_date", 
+                                        "days_to_election", "initial_weight", 
+                                        "weight", "pol_region")]
+
+kable(head(second_table, 5), col.names = names(second_table), 
+      booktabs = TRUE)
+
 ```
@@
 ```
+\newpage
 
@@
   labs(x = "Polling Method", y = "Number of Appearances", title = "Frequency of Polling Methods") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
 ```
+\newpage
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
 # Results {#sec-results}
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-10-21 scripts/03-clean_data.R
tracking which states are not being polled

@@ -18,6 +18,15 @@ raw_poll_data <- read_csv("data/raw_poll_data.csv")
 raw_poll_data <- raw_poll_data |>
   mutate(end_date = mdy(end_date)) 
 
+#!CHECKING STATES!#
+all_states <- state.name
+all_states <- c(all_states, "District of Columbia")
+states <- unique(raw_poll_data$state)
+
+missing_states <- setdiff(all_states, states)
+print(missing_states)
+#!CHECKING STATES!#
+
 raw_poll_data <- raw_poll_data |>
   filter(end_date >= as.Date("2024-07-22"))
 
@@ -154,6 +163,16 @@ region_mapping <- c(
 
 cleaned_poll_data <- cleaned_poll_data %>%
   mutate(pol_region = region_mapping[state])
+
+#!CHECKING STATES!#
+all_states <- state.name
+all_states <- c(all_states, "District of Columbia")
+states <- unique(cleaned_poll_data$state)
+
+missing_states <- setdiff(all_states, states)
+print(missing_states)
+#!CHECKING STATES!#
+
 
 View(cleaned_poll_data)
 str(cleaned_poll_data)

### 2024-10-21 scripts/06-model_data.R
added a bayesian model

@@ -136,6 +136,8 @@ for (region in pol_regions) {
 }
 
 states <- unique(poll_data$state)
+View(states)
+
 
 # Initialize a data frame to store state-level results
 state_results <- data.frame()
@@ -244,27 +246,227 @@ state_results_wide <- state_results_wide %>%
   )
 
 # Calculate total electoral votes for Harris
-harris_ev <- state_results_wide %>%
+harris_ev_lm <- state_results_wide %>%
   filter(winner == "Harris") %>%
   summarize(total_ev = sum(ev, na.rm = TRUE))
 
 # Calculate total electoral votes for Trump
-trump_ev <- state_results_wide %>%
+trump_ev_lm <- state_results_wide %>%
   filter(winner == "Trump") %>%
   summarize(total_ev = sum(ev, na.rm = TRUE))
 
 # Display the results
-print(harris_ev)
-print(trump_ev)
+print(harris_ev_lm)
+print(trump_ev_lm)
 
 
 
 
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+
+
+# Bayes Model
+#### Define State Priors ####
+# Define state categories based on political leaning
+state_priors <- data.frame(
+  state = c(
+    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
+    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia",
+    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
+    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
+    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
+    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
+    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina",
+    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
+    "Washington", "West Virginia", "Wisconsin", "Wyoming"
+  ),
+  category = c(
+    "Deep Red", "Deep Red", "Swing", "Deep Red", "Deep Blue", "Light Blue",
+    "Deep Blue", "Deep Blue", "Deep Blue", "Swing", "Light Red",
[…]

### 2024-10-22 paper/paper.qmd
added appendix

@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Model set-up
 
-## Model set-up
+We built three unique models, this allows us to choose the one that most closely fits the data historically, and weigh the pros and cons of each model to decide which one works best for us.
+This section will go through the process of how each of these models was developed and how we selected our final model.
+The first model was a conventional GLM, and the second was a Bayesian model with various differences and caveats.
+
+For the GLM model, we began by defining the 538 political regions [REF], and we fit a Harris and Trump GLM model for each of those regions.
+The intuition behind this is that voters in these regions are likely to vote in similar ways, therefore by pooling their preferences together we can get a more accurate model than one built at the State level.
+In other words, there would be too much noise in the State level model.
+Furthermore, some states have little to no polling data making the building of a model in these cases impossible (NOTE: how do we deal with these states later if there's no polling?).
+We then apply the polling data to these fitted regional models at the state level, to gauge Harris and Trump support.
+This allows us to predict the winner of each state, which informs us how many electoral college votes a candidate will get, and thus the winner of the election.
+
+The Bayesian model is necessarily more complicated.
+A major advantage of a Bayesian model is the ability to include prior information, encoded in the aptly named Prior.
+In the US elections states often vote very similarly [REF] across decades.
+We began to reflect this by classifying all states as either, Deep Red, Light Red, Swing, Light Blue, Dark Blue, to reflect their historical voting bias based on this list [REF].
+With the Priors encoded we fit a Harris and Trump model using Bayes R package [REF].
+We then synthesized a testing dataset for all the states, and used this data to get predictions of support for both candidates in each State.
+As was done in the previous model this gave us predictions on who would win each state, which allowed us to calculate the elector votes each candidate was likely to get and therefore the winner of the election.
 
@@
 
+MODEL NOT PERFECT WRITE AFTER DONE
+
+The GLM predicts Harris winning with [EV VOTES], with Trump only winning [EV VOTES].
+The Bayesinal model also predicts a Harris victory with [EV VOTES] and Trump with [EV VOTES].(Maybe better as a table).
+
 # Discussion {#sec-discussion}
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
 ## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
 
@@
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 
+One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
+Their analysis found 48% of participants favoured Harris while 50% favoured Trump.
+Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey.
+The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
+This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
+
+	One strength of this questionnaire is that information regarding the respondent’s confidence level and commitment level was also collected.
+With questions such as “How often do you make a point to read or listen to the news?” and “Are you certain to support that candidate, or do you think you may change your mind and support someone else?”, this poll is able to gain a more indepth view of how easily these participants may be swayed in the time between the polling and the real election (cite Fox article).
+As well, probabilistic statistical models based on past voting history, interest in current election, age, education, race, ethnicity, church attendance, and marital status were used to predict the respondents’ likeliness to vote; interestingly, results found that the results for the full sample vs subsample of 870 likely voters varied by a ±3%, which can be quite significant given how tight the race is currently (cite Fox article).
+
+	On the other hand, one potential weakness of this poll is that it was sponsored by Fox News, a news site that has historically been Republican-leaning.
+Though from the reported methodology alone, no evident biasing of the pollster can be observed, this potential source of biased reporting must be noted as it can affect the type of analyses that occurred and which key findings are the focus of news reports.
+As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article).
+
 ```

### 2024-10-22 paper/paper.qmd
updated appendix section

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "US Election 2024"
+subtitle: "Predicting election results"
 author: 
@@
 One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
-Their analysis found 48% of participants favoured Harris while 50% favoured Trump.
-Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey.
+Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phonecall interviews or an online survey.
 The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
 This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
+
+The key finding from this poll is that 48% of participants favoured Harris while 50% favoured Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters.
+Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote (cite Fox article).
+This report also notes that current results are Trump’s highest approval ratings since Biden dropped out of the race, while support for Harris is at its lowest.
 
@@
 As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article).
-
-```
+This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” (cite Fox article).
 

### 2024-10-25 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -26,3 +26,48 @@ max(simulated_polling_data$transparency_score) == 10
 min(simulated_polling_data$transparency_score) == 1
 
 
+#
+
+historical_wins <- tibble(
+  State = c("California", "New York", "Texas", "Florida"),
+  Party_Lean = c("Democrat", "Democrat", "Republican", "Republican")
+)
+
+
+# Function to check if simulated outcomes align with historical expectations
+test_that("State wins align with historical trends", {
+  # Filter Kamala Harris results
+  harris_results <- simulated_poll_data |>
+    filter(simulated_poll_data$`Candidate Favored` == "Harris")
+  
+  # Check if Harris is winning in historically Democrat-leaning states
+  historical_dem_states <- historical_wins %>%
+    filter(Party_Lean == "Democrat") %>%
+    pull(State)
+  
+  for (state in historical_dem_states) {
+    avg_support <- harris_results %>%
+      filter(State == state) %>%
+      summarize(avg_support = mean(Percentage, na.rm = TRUE)) %>%
+      pull(avg_support)
+    
+    # Test if average support in these states is reasonable for a Democrat
+
+        expect_true(avg_support > 50, info = paste("Kamala Harris is winning", state))
+  }
+  
+  # Check that she does not win in historically Republican states
+  historical_rep_states <- historical_wins %>%
+    filter(Party_Lean == "Republican") %>%
+    pull(State)
+  
+  for (state in historical_rep_states) {
+    avg_support <- harris_results %>%
+      filter(State == state) %>%
+      summarize(avg_support = mean(Percentage, na.rm = TRUE)) %>%
+      pull(avg_support)
+    
+    # Test if average support in these states is less than 50%
+    expect_true(avg_support <= 50, info = paste("Kamala Harris is losing", state))
+  }
+})

### 2024-10-25 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -26,21 +26,19 @@ max(simulated_polling_data$transparency_score) == 10
 min(simulated_polling_data$transparency_score) == 1
 
 
-#
+# Test historical wins
 
 historical_wins <- tibble(
   State = c("California", "New York", "Texas", "Florida"),
   Party_Lean = c("Democrat", "Democrat", "Republican", "Republican")
 )
 
-
-# Function to check if simulated outcomes align with historical expectations
 test_that("State wins align with historical trends", {
   # Filter Kamala Harris results
   harris_results <- simulated_poll_data |>
     filter(simulated_poll_data$`Candidate Favored` == "Harris")
   
-  # Check if Harris is winning in historically Democrat-leaning states
+  
   historical_dem_states <- historical_wins %>%
     filter(Party_Lean == "Democrat") %>%
     pull(State)
@@ -51,12 +49,12 @@ test_that("State wins align with historical trends", {
       summarize(avg_support = mean(Percentage, na.rm = TRUE)) %>%
       pull(avg_support)
     
-    # Test if average support in these states is reasonable for a Democrat
+    
 
         expect_true(avg_support > 50, info = paste("Kamala Harris is winning", state))
   }
   
-  # Check that she does not win in historically Republican states
+  
   historical_rep_states <- historical_wins %>%
     filter(Party_Lean == "Republican") %>%
     pull(State)
@@ -67,7 +65,7 @@ test_that("State wins align with historical trends", {
       summarize(avg_support = mean(Percentage, na.rm = TRUE)) %>%
       pull(avg_support)
     
-    # Test if average support in these states is less than 50%
+    
     expect_true(avg_support <= 50, info = paste("Kamala Harris is losing", state))
   }
 })

### 2024-10-25 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -25,3 +25,49 @@ clean_data$`Candidate Favored` |> class() == "character"
 max(cleaned_poll_data$transparency_score) == 10
 min(cleaned_poll_data$transparency_score) == 1
 
+
+# Test historical wins
+
+historical_wins <- tibble(
+  State = c("California", "Massachusetts", "Texas", "West Virigina"),
+  Party_Lean = c("Democrat", "Democrat", "Republican", "Republican")
+)
+
+
+
+test_that("State wins align with historical trends", {
+  # Filter Kamala Harris results
+  harris_results <- cleaned_poll_data |>
+    filter(cleaned_poll_data$candidate == "Harris")
+  
+  
+  historical_dem_states <- historical_wins %>%
+    filter(Party_Lean == "Democrat") %>%
+    pull(State)
+  
+  for (state in historical_dem_states) {
+    avg_support <- harris_results %>%
+      filter(state == state) %>%
+      summarize(avg_support = mean(support, na.rm = TRUE)) %>%
+      pull(avg_support)
+    
+    
+    
+    expect_true(avg_support > 50, info = paste("Kamala Harris is winning", state))
+  }
+  
+  # Check that she does not win in historically Republican states
+  historical_rep_states <- historical_wins %>%
+    filter(Party_Lean == "Republican") %>%
+    pull(State)
+  
+  for (state in historical_rep_states) {
+    avg_support <- harris_results %>%
+      filter(state == state) %>%
+      summarize(avg_support = mean(support, na.rm = TRUE)) %>%
+      pull(avg_support)
+    
+    # Test if average support in these states is less than 50%
+    expect_true(avg_support <= 50, info = paste("Kamala Harris is losing", state))
+  }
+})

### 2024-10-27 paper/paper.qmd
updated abstract

@@
 ---
-title: "US Election 2024"
-subtitle: "Predicting election results"
+title: "Forecasting the United States 2024 Presidential Election"
+subtitle: "Kamala Harris wins 60th presidency with 265 Electoral College Votes"
 author: 
@@
 date-format: long
-abstract: "Compiled polling surveys regarding the 2024 United States presidential election were analyzed in this report based on data from FiveThirtyEight [@538data].
+abstract: "This report analyzes compiled polling surveys for the 2024 U.S.
+presidential election using FiveThirtyEight data [@538data] and the 'pool of polls' method, which aggregates multiple polls to create a stable estimate of public opinion and reduce bias.
 From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system.
-The models forecast Kamala Harris as the winner of the US presidency.
+The models forecast Kamala Harris as the winner of election, winning 265 electoral college votes versus the 172 votes Donald Trump acquires.
 Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process.
@@
 Every four years, the United States presidential election occurs and is one of the most significant political events internationally.
-The complex process of the election involves both the popular and electoral college system, in which voters cast ballots on election day in each state.
+The complex process of the election involves both the popular and electoral college system.
 The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege].
@@
 The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege].
+
 Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs.
@@
 The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
-Pollster Name indicates which pollster the individual poll came from, which varies from individual companies, such as YouGov, to research institutions, such as Quinnipiac University and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time.
-This is to adjust for any bias.
+Pollster Name indicates which pollster the individual poll came from and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time to adjust for any bias.
 Method is the way the poll was deployed, including but not limited to live phone, online panel, and text-to-web.
-State indicates which state the poll was conducted.
 State indicates what state the poll was conducted and sample size is the size of the poll.
@@
 We also reorganized the dataset to include a poll region, which divides the US into geographical regions such as rust belt, northeast, etc.
-
-The table details the first few observations of the cleaned dataset, which we used to build our model.
-
-```{r}
-#| label: tbl-1
-#| tbl-cap: First few observations of cleaned data set
-#| echo: false
-
-first_table <- cleaned_poll_data[, c("pollster", "numeric_grade", "pollscore", 
-                                       "end_date", "transparency_score", 
-                                       "question_id", "method")]
-
-kable(head(first_table, 5), col.names = names(first_table), 
-      booktabs = TRUE)
-
-second_table <- cleaned_poll_data[, c("state", "sample_size", "candidate", 
-                                        "support", "election_date", 
-                                        "days_to_election", "initial_weight", 
-                                        "weight", "pol_region")]
-
-kable(head(second_table, 5), col.names = names(second_table), 
-      booktabs = TRUE)
-
-```
+The table in @sec-app-table details the first few observations of the cleaned dataset, which we used to build our model.
 
@@
 ```{r}
-#| label: fig-1
 #| fig-cap: Top 25 pollsters in overall polling data
@@
 
+## Analysis into a Pollster's Methodology
+
 One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
@@
 
-\newpage
+## Idealized Methodology and Survey
+
+## Table Detailing First Few Observations of Dataset {#sec-app-table}
+
+```{r}
+#| label: tbl-1
+#| tbl-cap: First few observations of cleaned data set
+#| echo: false
+
+first_table <- cleaned_poll_data[, c("pollster", "numeric_grade", "pollscore", 
+                                       "end_date", "transparency_score", 
+                                       "question_id", "method")]
+
+kable(head(first_table, 5), col.names = names(first_table), 
+      booktabs = TRUE)
+
+second_table <- cleaned_poll_data[, c("state", "sample_size", "candidate", 
+                                        "support", "election_date", 
+                                        "days_to_election", "initial_weight", 
+                                        "weight", "pol_region")]
+
+kable(head(second_table, 5), col.names = names(second_table), 
+      booktabs = TRUE)
+
+```
 

### 2024-10-28 paper/paper.qmd
Updated intro and estimand section

@@
 abstract: "This report analyzes compiled polling surveys for the 2024 U.S.
-presidential election using FiveThirtyEight data [@538data] and the 'pool of polls' method, which aggregates multiple polls to create a stable estimate of public opinion and reduce bias.
+presidential election using FiveThirtyEight data [@538data] and the 'pool of polls' method.
 From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system.
@@
 
-To forecast the 2024 United States presidential election, a generalized linear model and bayesian model were built.
-Our goal is to determine who will win the election based on the electoral college system.
-A key finding from our model is that we predict Kamala Harris to win the presidential election.
-This would be a historic win, as she would be the first woman to become US president.
-This would represent a significant milestone in breaking barriers related to gender and racial representation in the highest levels of American politics, symbolizing progress toward a more inclusive and diverse political landscape.
+To forecast the 2024 United States presidential election, a generalized linear model and Bayesian model were built.
+These models will help us predict who will win the election based on the electoral college system.
+A key finding from our model is that we predict Kamala Harris to win the presidential election with 265 electoral college votes.
+This would be a historic win, as she would be the first woman to become US president, a significant milestone in American politics.
 One drawback of our model stems from representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
@@
 The dataset used was obtained through FiveThirtyEight Interactives [@538data] and is focused on the United States presidential general polls for 2024.
-The polling data extracted are polling averages, where FiveThirtyEight [@538] includes individual polling data from all publicly available polls that meet their ethical and methodological standards and at minimum, must test Harris versus Trump, the two primary presidential nominees for each political party.
-From the individual polls, FiveThirtyEight [@538] will reweight and adjust the polls based on a few criteria, such as poll sample size, how recent it is, and house effects.
-The individual poll data is then added to the dataset, which is accessible for us to use.
+The polling data extracted utilize the "pool of polls" method, which combines multiple polling results into a single estimate.
+By aggregating data from several polls, this method aims to create a more stable and reliable measure of public opinion by reducing the impact of outliers, sampling error, and individual poll biases.
 This dataset is crucial to help us build a model to forecast the 2024 US Presidential election and can help us analyze key features in voter habits for future political analysis.
+Note that polling data collection stopped on October 19.
+Our model’s results are based solely on data available as of that date and do not incorporate any subsequent polling updates.
 
@@
 #| label: fig-2
-#| fig-cap: Frequency of Various Polling Methods
+#| fig-cap: Frequency of Various Polling Methods used by FiveThirtyEight in their "pool of polls"
 #| echo: false
@@
 #| label: fig-3
-#| fig-cap: Number of Votes per Candidate
+#| fig-cap: Number of Votes per Candidate. This reflects the popular vote for each candidate and does not account for the electoral college sytem. 
 #| echo: false
@@
 
-The key finding from this poll is that 48% of participants favoured Harris while 50% favoured Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters.
+The key finding from this poll is that 48% of participants favored Harris while 50% favored Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters.
 Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote (cite Fox article).
@@
 
+```{=latex}
+\newpage\newpage
+
+```
 # References

### 2024-10-28 paper/paper.qmd
updated a legend in graph

@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10), 
+        legend.position = "none")
 ```

### 2024-10-29 scripts/00-simulate_data.R
added TODOs

@@ -11,6 +11,10 @@ library(tidyverse)
 
 #### Simulate data #### 
 
+#TODO:
+#Make simulations reflect close to reality, for example Harris will definitly win Hawaii, and Trump 
+#will win Kansas
+
 # set up vectors where we will sample from
 states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
               "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 

### 2024-10-29 scripts/01-test_simulated_data.R

@@ -8,6 +8,7 @@
   # - access to cleaned datasets: access to and run 00-simulate_data.csv
   # - `tidyverse` library must be installed and loaded
 
+#TODO: Set up a more comprehensive set of tests. See Chapter 9 for all the tools that we need to use
 
 #### Workplace setup ####
 simulated_polling_data <- read_csv("data/00-simulate_data.csv")

### 2024-10-29 scripts/03-clean_data.R

@@ -11,6 +11,8 @@ library(tidyverse)
 library(lubridate)
 library(nanoparquet)
 
+#TODO Set up parquet data folder! (see chapter 10)
+
 #### Clean data ####
 # read file first
 raw_poll_data <- read_csv("data/raw_poll_data.csv")

### 2024-10-29 paper/paper.qmd
added TODOs

--- 2024-10-27 paper/paper.qmd
@@
 ---
-title: "Forecasting the United States 2024 Presidential Election"
-subtitle: "Kamala Harris wins 60th presidency with 265 Electoral College Votes"
+title: "My title"
+subtitle: "My subtitle if needed"
 author: 
@@
 date-format: long
-abstract: "This report analyzes compiled polling surveys for the 2024 U.S.
-presidential election using FiveThirtyEight data [@538data] and the 'pool of polls' method, which aggregates multiple polls to create a stable estimate of public opinion and reduce bias.
+abstract: "Compiled polling surveys regarding the 2024 United States presidential election were analyzed in this report based on data from FiveThirtyEight [@538data].
 From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system.
-The models forecast Kamala Harris as the winner of election, winning 265 electoral college votes versus the 172 votes Donald Trump acquires.
+The models forecast Kamala Harris as the winner of the US presidency.
 Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process.
@@
 Every four years, the United States presidential election occurs and is one of the most significant political events internationally.
-The complex process of the election involves both the popular and electoral college system.
+The complex process of the election involves both the popular and electoral college system, in which voters cast ballots on election day in each state.
 The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege].
@@
 The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege].
-
 Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs.
@@
 The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
-Pollster Name indicates which pollster the individual poll came from and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time to adjust for any bias.
+Pollster Name indicates which pollster the individual poll came from, which varies from individual companies, such as YouGov, to research institutions, such as Quinnipiac University and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time.
+This is to adjust for any bias.
 Method is the way the poll was deployed, including but not limited to live phone, online panel, and text-to-web.
+State indicates which state the poll was conducted.
 State indicates what state the poll was conducted and sample size is the size of the poll.
@@
 We also reorganized the dataset to include a poll region, which divides the US into geographical regions such as rust belt, northeast, etc.
-The table in @sec-app-table details the first few observations of the cleaned dataset, which we used to build our model.
+
+The table details the first few observations of the cleaned dataset, which we used to build our model.
+
+```{r}
+#| label: tbl-1
+#| tbl-cap: First few observations of cleaned data set
+#| echo: false
+
+first_table <- cleaned_poll_data[, c("pollster", "numeric_grade", "pollscore", 
+                                       "end_date", "transparency_score", 
+                                       "question_id", "method")]
+
+kable(head(first_table, 5), col.names = names(first_table), 
+      booktabs = TRUE)
+
+second_table <- cleaned_poll_data[, c("state", "sample_size", "candidate", 
+                                        "support", "election_date", 
+                                        "days_to_election", "initial_weight", 
+                                        "weight", "pol_region")]
+
+kable(head(second_table, 5), col.names = names(second_table), 
+      booktabs = TRUE)
+
+```
 
@@
 ```{r}
+#| label: fig-1
 #| fig-cap: Top 25 pollsters in overall polling data
@@
 
+TODO NEED TO JUSTIFY MY PRIORS MORE
+
 The Bayesian model is necessarily more complicated.
@@
 
+FOR IDEALIZED SURVEY
+!TODO like the other appendix, need to talk about depth about the techniques.
+Need to create a google form that is the actual survey.
+Every decision made about the idealized survey must be justified with literature.
+Every detail of the survey must be accounted for.
+Need to write about how the survey was created, and identify and justify all of the tradeoff made.
+TODO!
+
 # Appendix {#sec-appendices}
 
-## Analysis into a Pollster's Methodology
+!TODO need to add details about the techniques used by the pollster.
+For example, if stratified sampling was used, dive into what is stratified sampling.
+What is it good for why is works.
+Reference a academic paper about stratified sampling.
+TODO!
 
 One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
-Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phonecall interviews or an online survey.
+Their analysis found 48% of participants favoured Harris while 50% favoured Trump.
+Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey.
 The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
 This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
-
-The key finding from this poll is that 48% of participants favoured Harris while 50% favoured Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters.
-Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote (cite Fox article).
-This report also notes that current results are Trump’s highest approval ratings since Biden dropped out of the race, while support for Harris is at its lowest.
 
@@
 As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article).
-This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” (cite Fox article).
-
-## Idealized Methodology and Survey
-
-## Table Detailing First Few Observations of Dataset {#sec-app-table}
-
-```{r}
-#| label: tbl-1
-#| tbl-cap: First few observations of cleaned data set
-#| echo: false
-
-first_table <- cleaned_poll_data[, c("pollster", "numeric_grade", "pollscore", 
-                                       "end_date", "transparency_score", 
-                                       "question_id", "method")]
-
-kable(head(first_table, 5), col.names = names(first_table), 
-      booktabs = TRUE)
-
-second_table <- cleaned_poll_data[, c("state", "sample_size", "candidate", 
-                                        "support", "election_date", 
-                                        "days_to_election", "initial_weight", 
-                                        "weight", "pol_region")]
-
-kable(head(second_table, 5), col.names = names(second_table), 
-      booktabs = TRUE)
 
@@
 
+\newpage
+
 # References

### 2024-10-30 scripts/02-download_data.R
added historical voting history of each state

@@ -9,8 +9,18 @@
 #### Workplace Setup ####
 # install.packages("tidyverse")
 library(tidyverse)
+# Install devtools if you haven't already
+library(devtools)
+library(dplyr)
+
+# Install electionData package from GitHub
+devtools::install_github("jaytimm/PresElectionResults")
+
 
 #### Download data ####
+historical_state_data <- PresElectionResults::pres_by_state
+
+
 raw_poll_data <- 
   read_csv(
     file = 
@@ -20,6 +30,8 @@ raw_poll_data <-
   )
 
 
-View(raw_poll_data)
+#View(raw_poll_data)
+#View(historical_state_data)
 # save to csv
 write_csv(raw_poll_data, "data/raw_poll_data.csv")
+write_csv(historical_state_data, "data/historical_state_data.csv")

### 2024-10-30 scripts/03-clean_data.R

@@ -14,8 +14,13 @@ library(nanoparquet)
 #TODO Set up parquet data folder! (see chapter 10)
 
 #### Clean data ####
-# read file first
+
 raw_poll_data <- read_csv("data/raw_poll_data.csv")
+historical_state_data <- read_csv("data/historical_state_data.csv")
+
+# Mapping out the state abbreviations to their full names\
+cleaned_historical_state_data <- historical_state_data %>%
+  mutate(state_abbrev = state.name[match(state_abbrev, state.abb)])
 
 # filter out polls from before Biden dropped out of race
 raw_poll_data <- raw_poll_data |>
@@ -182,4 +187,6 @@ str(cleaned_poll_data)
 # write to csv and parquet
 write_csv(cleaned_poll_data, "data/cleaned_poll_data.csv")
 write_parquet(cleaned_poll_data, "data/cleaned_poll_data.parquet")
+write_csv(cleaned_historical_state_data, "data/cleaned_historical_data.csv")
+write_csv(cleaned_historical_state_data, "data/cleaned_historical_data.parquet")
 

### 2024-10-31 scripts/03-clean_data.R
fixed up

@@ -22,6 +22,9 @@ historical_state_data <- read_csv("data/historical_state_data.csv")
 cleaned_historical_state_data <- historical_state_data %>%
   mutate(state_abbrev = state.name[match(state_abbrev, state.abb)])
 
+cleaned_historical_state_data <- cleaned_historical_state_data |>
+  rename("state" = "state_abbrev")
+
 # filter out polls from before Biden dropped out of race
 raw_poll_data <- raw_poll_data |>
   mutate(end_date = mdy(end_date)) 
@@ -51,6 +54,7 @@ cleaned_poll_data <- cleaned_poll_data |>
          "support" = "pct")
 
 cleaned_poll_data <- na.omit(cleaned_poll_data)
+view(cleaned_poll_data)
 
 # Adding how many days before the election that the poll was conducted
 
@@ -181,7 +185,7 @@ missing_states <- setdiff(all_states, states)
 print(missing_states)
 #!CHECKING STATES!#
 
-
+View(cleaned_historical_state_data)
 View(cleaned_poll_data)
 str(cleaned_poll_data)
 # write to csv and parquet

### 2024-10-31 scripts/05-model_data.R

@@ -86,7 +86,7 @@ region_mapping <- c(
 )
 poll_data <- read_csv("data/cleaned_poll_data.csv")
 
-View(poll_data)
+#View(poll_data)
 
 #TODO add lm model, response: pct, predictors: everything else
 
@@ -230,6 +230,9 @@ electoral_votes <- data.frame(
     12, 4, 10, 3
   )
 )
+View(electoral_votes)
+
+write_csv(electoral_votes, "data/electoral_votes.csv")
 
 # Merge state_results with electoral votes
 state_results_ev <- merge(state_results, electoral_votes, by = "state")
@@ -258,213 +261,6 @@ trump_ev_lm <- state_results_wide %>%
 # Display the results
 print(harris_ev_lm)
 print(trump_ev_lm)
-
-
-
-
-
-
-
-# Bayes Model
-#### Define State Priors ####
-# Define state categories based on political leaning
-state_priors <- data.frame(
-  state = c(
-    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
-    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia",
-    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
-    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
-    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
-    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
-    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina",
-    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
-    "Washington", "West Virginia", "Wisconsin", "Wyoming"
-  ),
-  category = c(
-    "Deep Red", "Deep Red", "Swing", "Deep Red", "Deep Blue", "Light Blue",
-    "Deep Blue", "Deep Blue", "Deep Blue", "Swing", "Light Red",
-    "Deep Blue", "Deep Red", "Deep Blue", "Light Red", "Swing", "Deep Red", "Deep Red",
-    "Deep Red", "Light Blue", "Deep Blue", "Deep Blue", "Swing", "Light Blue",
[…]

### 2024-10-31 scripts/06-bayes_model.R

@@ -1,0 +1,202 @@
+#### Preamble ####
+# Purpose: Models the regional support, then calculates the EV each candidate is
+#          expected to get, using available data and priors where data is missing.
+# Author:  Robert Ford, Michelle Ji, Cher Ning
+# Date: 14 October 2024
+# Contact: robert.ford@mail.utoronto.ca, michelle.ji@mail.utoronto.ca, cher.ning@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+poll_data <- read_csv("data/cleaned_poll_data.csv")
+
+### Bayes Model ###
+
+# Define State Priors #
+# Define state categories based on political leaning
+state_priors <- data.frame(
+  state = c(
+    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
+    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia",
+    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
+    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
+    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
+    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
+    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina",
+    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
+    "Washington", "West Virginia", "Wisconsin", "Wyoming"
+  ),
+  category = c(
+    "Deep Red", "Deep Red", "Swing", "Deep Red", "Deep Blue", "Light Blue",
+    "Deep Blue", "Deep Blue", "Deep Blue", "Swing", "Light Red",
+    "Deep Blue", "Deep Red", "Deep Blue", "Light Red", "Swing", "Deep Red", "Deep Red",
+    "Deep Red", "Light Blue", "Deep Blue", "Deep Blue", "Swing", "Light Blue",
+    "Deep Red", "Light Red", "Light Red", "Deep Red", "Light Blue", "Light Blue",
+    "Deep Blue", "Light Blue", "Deep Blue", "Swing", "Deep Red",
[…]

### 2024-10-31 scripts/05-model_data.R
nm

@@ -262,6 +262,7 @@ trump_ev_lm <- state_results_wide %>%
 print(harris_ev_lm)
 print(trump_ev_lm)
 
+
 #### Save the Models (Optional) ####
 # Save the models
 saveRDS(model_harris, file = "models/model_harris_stan_glmer.rds")

### 2024-10-31 scripts/06-bayes_model.R
new bayes model

@@ -9,194 +9,393 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
-
+library(brms)
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
 #### Read data ####
 poll_data <- read_csv("data/cleaned_poll_data.csv")
-
+historical_state_data <- read_csv("data/cleaned_historical_data.csv")
+ev_votes <- read_csv("data/electoral_votes.csv")
 ### Bayes Model ###
+#View(historical_state_data)
+#View(poll_data)
 
-# Define State Priors #
-# Define state categories based on political leaning
-state_priors <- data.frame(
-  state = c(
-    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
-    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia",
-    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
-    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
-    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
-    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
-    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina",
-    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
-    "Washington", "West Virginia", "Wisconsin", "Wyoming"
-  ),
-  category = c(
-    "Deep Red", "Deep Red", "Swing", "Deep Red", "Deep Blue", "Light Blue",
-    "Deep Blue", "Deep Blue", "Deep Blue", "Swing", "Light Red",
-    "Deep Blue", "Deep Red", "Deep Blue", "Light Red", "Swing", "Deep Red", "Deep Red",
-    "Deep Red", "Light Blue", "Deep Blue", "Deep Blue", "Swing", "Light Blue",
-    "Deep Red", "Light Red", "Light Red", "Deep Red", "Light Blue", "Light Blue",
-    "Deep Blue", "Light Blue", "Deep Blue", "Swing", "Deep Red",
-    "Swing", "Deep Red", "Light Blue", "Swing", "Deep Blue", "Deep Red",
-    "Deep Red", "Deep Red", "Light Red", "Deep Red", "Deep Blue", "Light Blue",
[…]

### 2024-11-02 scripts/06-bayes_model.R
updated Baysian Model

@@ -1,401 +1,278 @@
 #### Preamble ####
-# Purpose: Models the regional support, then calculates the EV each candidate is
-#          expected to get, using available data and priors where data is missing.
+# Purpose: Models regional support and calculates the Electoral Votes (EV) each candidate is expected to get,
+#          using available polling data and historical averages where data is missing.
 # Author:  Robert Ford, Michelle Ji, Cher Ning
-# Date: 14 October 2024
-# Contact: robert.ford@mail.utoronto.ca, michelle.ji@mail.utoronto.ca, cher.ning@mail.utoronto.ca
+# Modified by: [Your Name]
+# Date: [Current Date]
+# Contact: [Your Contact Information]
 # License: MIT
+
+# -------------------------------------------------------------------------
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 library(brms)
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
+
 #### Read data ####
 poll_data <- read_csv("data/cleaned_poll_data.csv")
 historical_state_data <- read_csv("data/cleaned_historical_data.csv")
 ev_votes <- read_csv("data/electoral_votes.csv")
-### Bayes Model ###
-#View(historical_state_data)
-#View(poll_data)
 
+# Clean poll_data by removing rows with NAs in specified predictors
+poll_data <- poll_data %>%
+  drop_na(days_to_election, sample_size, transparency_score, pollscore, state)
 
-# Filter data from 2000 onwards
+# Filter historical data from 2000 onwards
 historical_recent <- historical_state_data %>%
   filter(year >= 2000)
 
 # Calculate average support for each party in each state
-state_averages <- historical_recent %>%
+state_stats <- historical_recent %>%
   group_by(state) %>%
   summarize(
     avg_democrat = mean(democrat, na.rm = TRUE),
     avg_republican = mean(republican, na.rm = TRUE)
   )
 
-# Filter poll data for Trump and Harris
-poll_data_trump <- poll_data %>% filter(candidate == "Trump")
-poll_data_harris <- poll_data %>% filter(candidate == "Harris")
+# Separate polling data for Trump and Harris
+poll_data_trump <- poll_data %>%
+  filter(candidate == "Trump") %>%
+  mutate(state = as.character(state))  # Ensure 'state' is character
 
-# Define the model formula
-formula <- support ~ days_to_election + sample_size + transparency_score + pollscore
+poll_data_harris <- poll_data %>%
+  filter(candidate == "Harris") %>%
[…]

### 2024-11-02 scripts/05-model_data.R
fixed states list

@@ -136,7 +136,7 @@ for (region in pol_regions) {
 }
 
 states <- unique(poll_data$state)
-View(states)
+#View(states)
 
 
 # Initialize a data frame to store state-level results
@@ -209,7 +209,7 @@ View(state_results)
 electoral_votes <- data.frame(
   state = c(
     "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
-    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia",
+    "Connecticut", "Delaware", "Florida", "Georgia",
     "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
     "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
     "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
@@ -220,7 +220,7 @@ electoral_votes <- data.frame(
   ),
   ev = c(
     9, 3, 11, 6, 54, 10,
-    7, 3, 3, 30, 16,
+    7, 3, 30, 16,
     4, 4, 19, 11, 6, 6, 8,
     8, 4, 10, 11, 15, 10,
     6, 10, 4, 5, 6, 4,

### 2024-11-02 scripts/03-clean_data.R
added checks for data cleaning

@@ -97,6 +97,7 @@ cleaned_poll_data$state[cleaned_poll_data$state == "Maine CD-1"] <- "Maine"
 cleaned_poll_data$state[cleaned_poll_data$state == "Maine CD-2"] <- "Maine"
 cleaned_poll_data$state[cleaned_poll_data$state == "Nebraska CD-1"] <- "Nebraska"
 cleaned_poll_data$state[cleaned_poll_data$state == "Nebraska CD-2"] <- "Nebraska"
+cleaned_poll_data$state[cleaned_poll_data$state == "Nebraska CD-3"] <- "Nebraska"
 
 
 # Define the state to region mapping
@@ -184,6 +185,19 @@ states <- unique(cleaned_poll_data$state)
 missing_states <- setdiff(all_states, states)
 print(missing_states)
 #!CHECKING STATES!#
+
+
+# Summarize missing values for each predictor
+missing <- poll_data %>%
+  summarize(
+    missing_days_to_election = sum(is.na(days_to_election)),
+    missing_sample_size = sum(is.na(sample_size)),
+    missing_transparency_score = sum(is.na(transparency_score)),
+    missing_pollscore = sum(is.na(pollscore)),
+    missing_state = sum(is.na(state))
+  )
+
+View(missing)
 
 View(cleaned_historical_state_data)
 View(cleaned_poll_data)

### 2024-11-02 scripts/02-download_data.R
added last date that data was collected

@@ -6,6 +6,8 @@
 # License: MIT
 # Prerequisites: none
 
+#LAST RUN 31 October
+
 #### Workplace Setup ####
 # install.packages("tidyverse")
 library(tidyverse)

### 2024-11-02 scripts/06-bayes_model.R
added diagnostics testing

@@ -56,6 +56,7 @@ states_with_poll <- union(states_with_poll_trump, states_with_poll_harris)
 
 # States with no polling data
 states_without_poll <- setdiff(all_states, states_with_poll)
+print(states_without_poll)
 
 #### Merge state_stats with Electoral Votes ####
 
@@ -276,3 +277,66 @@ avg_harris_ev <- mean(harris_ev_counts)
 
 cat("Average Electoral Votes for Trump:", round(avg_trump_ev, 2), "\n")
 cat("Average Electoral Votes for Harris:", round(avg_harris_ev, 2), "\n")
+
+#### Save the Models ####
+saveRDS(model_harris, file = "models/bayes_model_harris.rds")
+saveRDS(model_trump, file = "models/bayes_model_trump.rds")
+
+### Model Diagnostics ###
+
+# Model Convergence
+
+# Check convergence for Trump model
+print(summary(model_trump)$coefficients)
+print(rhat(model_trump))  # Should be ~1
+
+# Check convergence for Harris model
+print(summary(model_harris)$coefficients)
+print(rhat(model_harris))  # Should be ~1
+
+# Plot trace plots for Trump
+plot(model_trump, plotfun = "trace")
+
+# Plot trace plots for Harris
+plot(model_harris, plotfun = "trace")
+
+# PPC Checks
+
+# Enhanced PPC with different types
+library(bayesplot)
+
+# Trump PPC
+ppc_dens_overlay(y = poll_data_trump$support, 
+                 yrep = posterior_predict(model_trump)) +
+  ggtitle("Posterior Predictive Density Overlay - Trump")
+
+# Harris PPC
+ppc_dens_overlay(y = poll_data_harris$support, 
+                 yrep = posterior_predict(model_harris)) +
+  ggtitle("Posterior Predictive Density Overlay - Harris")
+
+# PPC for residuals
+pp_check(model_trump, type = "residuals")
+pp_check(model_harris, type = "residuals")
+
+# Residual Analysis
+
+# Function to plot residuals
+plot_residuals <- function(model, candidate_name) {
+  residuals <- residuals(model)
+  fitted <- fitted(model)
+  
+  ggplot(data = data.frame(fitted, residuals), aes(x = fitted, y = residuals)) +
+    geom_point(alpha = 0.5) +
+    geom_hline(yintercept = 0, linetype = "dashed") +
+    labs(title = paste("Residuals vs Fitted -", candidate_name),
[…]

### 2024-11-03 scripts/06-bayes_model.R
added more graphs

@@ -13,6 +13,9 @@
 library(tidyverse)
 library(rstanarm)
 library(brms)
+library(usmap)      # For US maps
+library(ggplot2)
+
 
 #### Read data ####
 poll_data <- read_csv("data/cleaned_poll_data.csv")
@@ -340,3 +343,188 @@ plot_residuals(model_trump, "Trump")
 
 # Plot residuals for Harris
 plot_residuals(model_harris, "Harris")
+
+
+
+# Graphing
+
+#### Graph 1: Probability of Victory Over Time ####
+
+# Define a range of days_to_election (e.g., from 100 to 0)
+days_range <- seq(0, 100, by = 1)  # Adjust the range as needed
+
+# Function to generate newdata for a specific days_to_election
+generate_newdata_time <- function(days) {
+  # Assume other predictors are held constant at their mean values
+  newdata_trump_time <- tibble(
+    days_to_election = days,
+    sample_size = mean(poll_data_trump$sample_size, na.rm = TRUE),
+    transparency_score = mean(poll_data_trump$transparency_score, na.rm = TRUE),
+    pollscore = mean(poll_data_trump$pollscore, na.rm = TRUE),
+    state = "California"  # Placeholder; state effect can be averaged or set to a reference
+  )
+  
+  newdata_harris_time <- tibble(
+    days_to_election = days,
+    sample_size = mean(poll_data_harris$sample_size, na.rm = TRUE),
+    transparency_score = mean(poll_data_harris$transparency_score, na.rm = TRUE),
+    pollscore = mean(poll_data_harris$pollscore, na.rm = TRUE),
+    state = "California"  # Placeholder; state effect can be averaged or set to a reference
+  )
+  
+  return(list(trump = newdata_trump_time, harris = newdata_harris_time))
+}
+
+# Initialize vectors to store probabilities
+prob_trump_victory <- numeric(length(days_range))
+prob_harris_victory <- numeric(length(days_range))
+
+# Loop over each day and calculate victory probabilities
+for (i in seq_along(days_range)) {
+  days <- days_range[i]
+  
+  # Generate newdata for the current day
+  newdata_list <- generate_newdata_time(days)
+  newdata_trump <- newdata_list$trump
+  newdata_harris <- newdata_list$harris
+  
+  # Predict support using the models
+  if (!is.null(model_trump)) {
+    pred_trump <- posterior_predict(model_trump, newdata = newdata_trump, draws = 1000)
+  } else {
[…]

### 2024-11-03 scripts/03-clean_data.R
commented View so the file runs without opening a bunch of df's

@@ -54,7 +54,7 @@ cleaned_poll_data <- cleaned_poll_data |>
          "support" = "pct")
 
 cleaned_poll_data <- na.omit(cleaned_poll_data)
-view(cleaned_poll_data)
+#view(cleaned_poll_data)
 
 # Adding how many days before the election that the poll was conducted
 
@@ -197,11 +197,11 @@ missing <- poll_data %>%
     missing_state = sum(is.na(state))
   )
 
-View(missing)
+#View(missing)
 
-View(cleaned_historical_state_data)
-View(cleaned_poll_data)
-str(cleaned_poll_data)
+#View(cleaned_historical_state_data)
+#View(cleaned_poll_data)
+#str(cleaned_poll_data)
 # write to csv and parquet
 write_csv(cleaned_poll_data, "data/cleaned_poll_data.csv")
 write_parquet(cleaned_poll_data, "data/cleaned_poll_data.parquet")

### 2024-11-03 paper/paper.qmd
updated paper

--- 2024-10-28 paper/paper.qmd
@@
 
-## Model set-up
+To predict state level support and calculate the expected Electoral Votes (EV) for each candidate in the upcoming US election, we employed a Bayesian linear regression model.
+The model is mathematically specified as follows:
 
-We built three unique models, this allows us to choose the one that most closely fits the data historically, and weigh the pros and cons of each model to decide which one works best for us.
-This section will go through the process of how each of these models was developed and how we selected our final model.
-The first model was a conventional GLM, and the second was a Bayesian model with various differences and caveats.
+$$
+\text{Support}_i = \beta_0 + \beta_1 \times \text{Sample Size}_i + \beta_2 \times \text{Days to election}_i + \beta_3 \times \text{Transparency Score}_i + \beta_4 \times \text{Pollscore}_i + \gamma \times \text{State}_i + \epsilon_i
+$$
 
-For the GLM model, we began by defining the 538 political regions [REF], and we fit a Harris and Trump GLM model for each of those regions.
-The intuition behind this is that voters in these regions are likely to vote in similar ways, therefore by pooling their preferences together we can get a more accurate model than one built at the State level.
-In other words, there would be too much noise in the State level model.
-Furthermore, some states have little to no polling data making the building of a model in these cases impossible (NOTE: how do we deal with these states later if there's no polling?).
-We then apply the polling data to these fitted regional models at the state level, to gauge Harris and Trump support.
-This allows us to predict the winner of each state, which informs us how many electoral college votes a candidate will get, and thus the winner of the election.
+- **support<sub>i</sub>**  
+  *Definition:* The support level for candidate \( i \) in a given poll.  
+  *Justification:* Measures the proportion of respondents supporting candidate \( i \), providing insight into their current standing.
 
-The Bayesian model is necessarily more complicated.
-A major advantage of a Bayesian model is the ability to include prior information, encoded in the aptly named Prior.
-In the US elections states often vote very similarly [REF] across decades.
-We began to reflect this by classifying all states as either, Deep Red, Light Red, Swing, Light Blue, Dark Blue, to reflect their historical voting bias based on this list [REF].
-With the Priors encoded we fit a Harris and Trump model using Bayes R package [REF].
-We then synthesized a testing dataset for all the states, and used this data to get predictions of support for both candidates in each State.
-As was done in the previous model this gave us predictions on who would win each state, which allowed us to calculate the elector votes each candidate was likely to get and therefore the winner of the election.
+- **sample\_size<sub>i</sub>**  
+  *Definition:* Represents the number of respondents in poll \( i \).  
+  *Justification:* Larger sample sizes generally yield more accurate estimates of support, reducing sampling variability.
 
-### Model justification
+- **days\_to\_election<sub>i</sub>**  
+  *Definition:* Denotes the number of days remaining until the election when poll \( i \) was conducted.  
+  *Justification:* Accounts for temporal proximity to the election, capturing potential fluctuations in support as the election approaches.
+
+- **transparency\_score<sub>i</sub>**  
+  *Definition:* A numerical score reflecting the pollster’s transparency regarding their methodology, ranging up to 10, with higher scores indicating greater transparency.  
+  *Justification:* Higher transparency scores may correlate with the reliability and credibility of poll results, influencing voter trust.
+
+- **pollscore<sub>i</sub>**  
+  *Definition:* A numeric value representing the reliability and bias of the pollster, where negative values denote better reliability.  
+  *Justification:* Accounts for systematic deviations in poll results based on pollster performance, ensuring more accurate support estimates.
+
+- **state<sub>i</sub>**  
+  *Definition:* Captures state-specific effects as fixed effects.  
+  *Justification:* Allows the model to account for regional variations in support that are not explained by other predictors, incorporating state-specific nuances.
+
+- **ϵ<sub>i</sub>**  
+  *Definition:* The error term, assumed to follow a normal distribution.  
+  *Justification:* Represents unexplained variability in the model, ensuring that the residuals meet the assumptions of the statistical analysis.
+
+Variable Definitions:
+
+- **Sample Size (\(\text{sample\_size}\))**: The size of the poll is included as larger samples generally provide more accurate estimates of support, reducing sampling variability.
+
+- **Days to Election (\(\text{days\_to\_election}\))**: This variable accounts for the temporal proximity to the election, as support levels may fluctuate as the election approaches, capturing trends in voter sentiment over time.
+
+- **Transparency Score (\(\text{transparency\_score}\))**: A higher transparency score indicates greater disclosure of poll methodology, potentially correlating with the reliability and credibility of the poll results.
+
+- **Pollscore (\(\text{pollscore}\))**: This score represents the inherent bias and error of the pollster, with lower (more negative) values indicating higher reliability.
+It accounts for systematic deviations in poll results based on historical pollster performance [REF 538 METHOD].
+
+- **State (\(\text{state}\))**: Including state as a fixed effect allows the model to account for regional variations in support that are not captured by other predictors, ensuring state-specific nuances are incorporated into the support estimates.
+
+An alternative approach considered was modeling State as a random effect to account for unobserved heterogeneity across states.
+However, given the focus on specific state-level predictions and the availability of sufficient data per state, fixed effects are chosen in this case.
+Ideally we would fit both models and use diagnostics to choose which has lower bais and variance, and thus better predictive quality.
+Additionally, simpler models excluding variables like transparency_score and pollscore were evaluated but found inadequate in capturing the state-by-state differences in the transparency and reliability of the polls, potentially leading to biased support estimates.
+
+Process:
+
+For each candidate and state, we aggregated polling data by calculating the mean values of each predictor.
+For example, here is the mathematical notation of Mean Days to Election, this value was calculated for each state for both candidates:
+
+$\text{Mean Days to Election}_s = \frac{1}{N_s} \sum_{i=1}^{N_s} \text{Days to Election}_i$
+
+where \(N_s\) is the number of polls for state \(s\).
+
+Averaging predictors per state smooths out poll-to-poll variability and provides a representative set of predictors for each state, allowing use to make reliable state-level predictions.
+Instead of averaging, a training/testing split or regional aggregations based on 538's political regions could have been employed.
+However, a training/testing split is beyond the scope of this paper, and using political regions would have complicated the derivation of the model.
+
+Using the averaged state level predictors, we generated posterior predictions by drawing 1,000 samples from the posterior distributions of the model parameters.
+These predictions simulate 1,000 possible election outcomes, and by taking the proportion of victories for Harris and Trump we estimate the probability each candidate has of winning the election.
+To generate an election map, we simply take the average outcome of each state across the 1,000 simulations.
+Alternative methods could include using point estimates from the posterior mean; however, simulating multiple outcomes offers a more comprehensive assessment of uncertainty and variability in election outcomes.
+
+For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
+This decision is based on the assumption that certain states are very unlikely to change their voting patterns, reducing the necessity for current polling data.
+The ten states that do not have polling data [REF] have not changed parties in at least the last 20 plus years.
+This may be why they do not have polling data, as it would be a waste of resources for agencies to poll them.
+Alternatively, imputation methods or political regional averages could have been used, but historical averages are used for their simplicity and relevance.
+
+The Bayesian models were implemented using the `rstanarm` package in R, which interfaces with Stan for efficient Markov Chain Monte Carlo (MCMC) sampling.
+Data manipulation and visualization were conducted using the `tidyverse` suite of packages, while model diagnostics leveraged `bayesplot` and other related packages.
+This combination of tools facilitated a streamlined workflow for model fitting, prediction, and validation.
+
+-Diagnostics
+
+Comprehensive diagnostics were performed to assess model convergence, fit, and predictive performance.
+Graphs and more details can be found in the relevant appendices.
+We evaluated model convergence using the R-hat statistic and trace plots.
+All R-hat values for model parameters were below 1.1, indicating satisfactory convergence.
+Trace plots revealed well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process [REF].
+
+Posterior predictive checks were conducted to assess the model's ability to replicate observed data.
+Density overlay plots for both Trump and Harris models indicated reasonable fits, although some discrepancies were noted near the peaks of the distributions.
+Specifically, the Trump model's replicated data tended to be slightly smaller around the peak, while the Harris model exhibited more extreme differences at the peak.
+Residual plots appeared randomly scattered, suggesting no major systematic biases, although some clumping was observed, indicating areas where the model may fit less effectively [REF].
+
+Residual plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines.
+This pattern may indicate unmodeled heterogeneity or data limitations, suggesting areas where the model's fit could be improved.
+Despite these observations, the overall residual distribution did not reveal significant systematic errors, supporting the model's general adequacy.
+
+Ideally several models would be created then using model selection techniques would be used to determine the best model for predicting the outcome of the US election.
+Due to time constraints the folling models were considered but not implemented.
+A set of Bayesian models with state as a random effect and models excluding specific predictors like transparency_score.
+More complex models with additional interaction terms or non-linear components were also considered
+
+#### Assumptions and Limitations
+
+The model operates under several key assumptions:
+
+1. **Linearity:** The relationship between predictors and support is linear.
+2. **Normality of Residuals:** The error terms follow a normal distribution.
+3. **Independence:** Observations are independent given the predictors.
+4. **Stable Partisan Leanings:** Historical averages adequately represent states without current polling data.
+
+**Limitations:**
+
+- **Model Misspecification:** If non-linear relationships exist between predictors and support, the linear model may not capture these dynamics effectively.
+- **Reliance on Historical Data:** Using historical averages for certain states may not account for recent political shifts or emerging trends.
+- **Residual Clumping:** Observed clumping in residual plots suggests potential areas for model refinement, such as incorporating additional predictors or interaction terms.
+
+The final Bayesian linear regression model balances complexity and interpretability, incorporating relevant predictors to capture poll reliability and state-specific effects while maintaining computational efficiency.
+The inclusion of transparency_score and pollscore enhances the model's ability to account for poll quality, while fixed state effects ensure accurate regional support estimates.
+Validation through convergence diagnostics, posterior predictive checks, and residual analysis confirms the model's robustness and reliability within its assumptions.
+
+All analyses were conducted using R (version 4.2.1) with the following key packages:
+
+- **tidyverse:** For data manipulation and visualization.
+- **rstanarm:** For Bayesian model fitting using Stan.
+- **brms:** Alternative package for Bayesian regression modeling.
+- **bayesplot:** For comprehensive model diagnostics and visualization.
 
 # Results {#sec-results}
-
-MODEL NOT PERFECT WRITE AFTER DONE
-
-The GLM predicts Harris winning with [EV VOTES], with Trump only winning [EV VOTES].
-The Bayesinal model also predicts a Harris victory with [EV VOTES] and Trump with [EV VOTES].(Maybe better as a table).
 
@@
 
+FOR IDEALIZED SURVEY
+!TODO like the other appendix, need to talk about depth about the techniques.
+Need to create a google form that is the actual survey.
+Every decision made about the idealized survey must be justified with literature.
+Every detail of the survey must be accounted for.
+Need to write about how the survey was created, and identify and justify all of the tradeoff made.
+TODO!
+
 # Appendix {#sec-appendices}
 
+!TODO need to add details about the techniques used by the pollster.
+For example, if stratified sampling was used, dive into what is stratified sampling.
+What is it good for why is works.
+Reference a academic paper about stratified sampling.
+TODO!
+
+One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
+Their analysis found 48% of participants favoured Harris while 50% favoured Trump.
+Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey.
+The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
+This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
+=======
 ## Analysis into a Pollster's Methodology

### 2024-11-03 paper/paper.qmd

@@
 
-## First discussion point {#sec-first-point}
+## Implications and Potential Weaknesses {#sec-disc-imp}
 
-## Second discussion point
+Our model predicting the U.S.
+election carries important social and political implications.
+Accurate forecasts can shape voter perceptions, potentially affecting turnout by leading some individuals to feel their votes are less crucial if a particular outcome seems likely.
+Additionally, these predictions inform campaign strategies, prompting candidates to focus resources on undecided voters in critical swing states, which play a pivotal role in tight elections like the upcoming 2024 race.
+For instance, our model anticipates that key swing states such as Pennsylvania, Michigan, and Wisconsin will award their electoral votes to Kamala Harris, contributing to her victory.
+However, the forecast indicates that Harris is expected to win by narrow margins of approximately 1.98%, 1.47%, and 1.46% in these states, respectively.
 
-## Third discussion point
+Our model offers valuable insights into the landscape of the current U.S.
+election; however, it does face a few limitations.
+One of the primary challenges stems from gaps in our dataset, specifically the absence of polling data for several states.
+When the polling data was collected and compiled by @538, it did not comprehensively include every state, leading to missing data points that affect our predictions.
+Our model relies on the electoral college system, where each state’s popular vote determines its allocation of electoral votes, ultimately deciding the election’s outcome.
+Missing data from certain states introduces an incomplete representation of the voting scene, affecting the accuracy of our predictions.
+Each state often exhibits unique voting behaviors and excluding certain states can introduce a bias in our model.
+This is especially problematic if the omitted states are swing states or tend to have historically unpredictable voting patterns, as these states can play a decisive role in election outcomes.
+Moreover, missing data for larger states further exacerbates the issue by potentially leading to a miscalculation of the Electoral College outcome.
+States with significant electoral votes carry substantial weight, and without accurate polling information from them, our model may struggle to forecast the election winner effectively.
 
-## Weaknesses and next steps
+Another limitation of our model is that we stopped pulling new poll data from @538 as of October 19, 2024.
+Consequently, the model does not account for any last-minute events that could have impacted voter opinions from then until election day.
+This could potentially lead to misleading predictions if there has been a significant change in public sentiment.
+Furthermore, last-minute shifts in public opinion could result in different voter turnout than initially anticipated or introduce polling response biases, which predictive models may struggle to capture if they are based primarily on earlier data.
+
+## Next Steps {#sec-disc-next}
+
+Consequently, our forecast suggests that candidates should concentrate their efforts on swing states, as increased campaigning in these areas could significantly impact voter turnout and the popular vote.
+Additionally, enlisting high-profile figures or celebrities, like Elon Musk or LeBron James—who have recently endorsed Trump and Harris, respectively—could enhance their appeal in these swing states.
+These endorsements would help candidates reach diverse and niche demographics, broadening their voter base and potentially swaying undecided voters.
+
+## Betting Markets {#sec-disc-bet}
 
@@
 
-One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
-Their analysis found 48% of participants favoured Harris while 50% favoured Trump.
-Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey.
-The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
-This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
 =======
@@
 
-One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article).
+One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research [@fox].
 Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phonecall interviews or an online survey.
-The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article).
-This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
+The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers [@fox].
+This sampling method is a type of probability sampling in which each unit's chance of being included is dependent upon their size, a measure that is based off of some characteristic that is known about every single unit and often related to the main variable of interest.
+By giving a higher probability of inclusion to units of greater importance based on its size during sampling, a more accurate estimate can be obtained [@Latpate2021].
+In the case of this poll, the number of voters per state region was used to determine the proportion of individuals contacted [@fox].
+This means a greater number of residents in larger states, such as California, would be contacted compared to smaller states, such as Hawaii.
+This method of sampling ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
 
 The key finding from this poll is that 48% of participants favored Harris while 50% favored Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters.
-Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote (cite Fox article).
+Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote  [@fox].
 This report also notes that current results are Trump’s highest approval ratings since Biden dropped out of the race, while support for Harris is at its lowest.
@@
 One strength of this questionnaire is that information regarding the respondent’s confidence level and commitment level was also collected.
-With questions such as “How often do you make a point to read or listen to the news?” and “Are you certain to support that candidate, or do you think you may change your mind and support someone else?”, this poll is able to gain a more indepth view of how easily these participants may be swayed in the time between the polling and the real election (cite Fox article).
-As well, probabilistic statistical models based on past voting history, interest in current election, age, education, race, ethnicity, church attendance, and marital status were used to predict the respondents’ likeliness to vote; interestingly, results found that the results for the full sample vs subsample of 870 likely voters varied by a ±3%, which can be quite significant given how tight the race is currently (cite Fox article).
+With questions such as “How often do you make a point to read or listen to the news?” and “Are you certain to support that candidate, or do you think you may change your mind and support someone else?”, this poll is able to gain a more indepth view of how easily these participants may be swayed in the time between the polling and the real election  [@fox].
+As well, probabilistic statistical models based on past voting history, interest in current election, age, education, race, ethnicity, church attendance, and marital status were used to predict the respondents’ likeliness to vote; interestingly, results found that the results for the full sample vs subsample of 870 likely voters varied by a ±3%, which can be quite significant given how tight the race is currently  [@fox].
 
@@
 Though from the reported methodology alone, no evident biasing of the pollster can be observed, this potential source of biased reporting must be noted as it can affect the type of analyses that occurred and which key findings are the focus of news reports.
-As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article).
-This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” (cite Fox article).
+As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis  [@fox].
+This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” [@fox].
 

### 2024-11-03 scripts/06-bayes_model.R
removed not working graph

@@ -290,11 +290,9 @@ saveRDS(model_trump, file = "models/bayes_model_trump.rds")
 # Model Convergence
 
 # Check convergence for Trump model
-print(summary(model_trump)$coefficients)
 print(rhat(model_trump))  # Should be ~1
 
 # Check convergence for Harris model
-print(summary(model_harris)$coefficients)
 print(rhat(model_harris))  # Should be ~1
 
 # Plot trace plots for Trump
@@ -348,90 +346,10 @@ plot_residuals(model_harris, "Harris")
 
 # Graphing
 
-#### Graph 1: Probability of Victory Over Time ####
-
-# Define a range of days_to_election (e.g., from 100 to 0)
-days_range <- seq(0, 100, by = 1)  # Adjust the range as needed
-
-# Function to generate newdata for a specific days_to_election
-generate_newdata_time <- function(days) {
-  # Assume other predictors are held constant at their mean values
-  newdata_trump_time <- tibble(
-    days_to_election = days,
-    sample_size = mean(poll_data_trump$sample_size, na.rm = TRUE),
-    transparency_score = mean(poll_data_trump$transparency_score, na.rm = TRUE),
-    pollscore = mean(poll_data_trump$pollscore, na.rm = TRUE),
-    state = "California"  # Placeholder; state effect can be averaged or set to a reference
-  )
-  
-  newdata_harris_time <- tibble(
-    days_to_election = days,
-    sample_size = mean(poll_data_harris$sample_size, na.rm = TRUE),
-    transparency_score = mean(poll_data_harris$transparency_score, na.rm = TRUE),
-    pollscore = mean(poll_data_harris$pollscore, na.rm = TRUE),
-    state = "California"  # Placeholder; state effect can be averaged or set to a reference
-  )
-  
-  return(list(trump = newdata_trump_time, harris = newdata_harris_time))
-}
-
 # Initialize vectors to store probabilities
 prob_trump_victory <- numeric(length(days_range))
 prob_harris_victory <- numeric(length(days_range))
 
-# Loop over each day and calculate victory probabilities
-for (i in seq_along(days_range)) {
-  days <- days_range[i]
-  
-  # Generate newdata for the current day
-  newdata_list <- generate_newdata_time(days)
-  newdata_trump <- newdata_list$trump
-  newdata_harris <- newdata_list$harris
-  
-  # Predict support using the models
-  if (!is.null(model_trump)) {
[…]

### 2024-11-03 paper/paper.qmd
added results section

@@
 ---
+![State Election Predictions](other/plots/election_map.png){width=80%}
 
@@
 library(kableExtra)
+library(usmap)      # For US maps
+library(ggplot2)
+
+plot_path <- here::here("other", "plots", "election_map.png")
 
@@
 
+Our model predicts that Harris has a 63.8% chance of winning the US presidential election.
+While Trump only has a 36.2% chance of winning.
+On average across the 1,000 simulations Harris wins 275.73 Electoral Votes, while Trump wins on average 259.27 votes.
+
+```{r}
+election_results %>%
+  kable("html", caption = "Election Prediction Results") %>%
+  kable_styling(
+    "bordered",
+    full_width = FALSE,
+    position = "center"
+  ) %>%
+  add_header_above(c(" " = 1, "Candidates" = 2)) %>%
+  row_spec(0, bold = TRUE) # Makes the header row bold
+
+# Display a nicely formatted table
+kable(election_results, caption = "Election Prediction Results")
+
+```
+
+Based on the simulated elections, this map illustrates the most likely winner of each state. In this scenario Harris wins 277 Electoral Votes, and Trump wins 262.[Appendix]
+
+```{r}
+knitr::include_graphics(plot_path)
+
+```
+
 # Discussion {#sec-discussion}

### 2024-11-03 scripts/06-bayes_model.R
added support for election map

@@ -224,12 +224,6 @@ for (state in states_without_poll) {
   harris_support_matrix[, state] <- avg_democrat
 }
 
-# Verify the support matrices
-print(dim(trump_support_matrix))  # Should be 1000 x number of states
-print(dim(harris_support_matrix)) # Should be 1000 x number of states
-print(head(trump_support_matrix))
-print(head(harris_support_matrix))
-
 #### Step 5: Simulate 1000 Elections ####
 
 # Initialize vectors to store EV counts for each simulation
@@ -420,7 +414,7 @@ candidate_colors <- c("Trump" = "#FF0000",   # Red
                       "Tie" = "#808080")      # Gray
 
 # Plot the map
-plot_usmap(data = map_data_combined, regions = "states", values = "winner") +
+election_map <- plot_usmap(data = map_data_combined, regions = "states", values = "winner") +
   scale_fill_manual(
     values = candidate_colors,
     name = "Winning Candidate",
@@ -428,10 +422,8 @@ plot_usmap(data = map_data_combined, regions = "states", values = "winner") +
   ) +
   labs(
     title = "State Election Predictions",
-    subtitle = "Based on 1000 Simulated Elections",
-    caption = "Red: Trump | Blue: Harris | Gray: Tie"
   ) +
-  theme_minimal() +
+  theme_void() +
   theme(
     plot.title = element_text(size = 20, hjust = 0.5),
     plot.subtitle = element_text(size = 14, hjust = 0.5),
@@ -439,6 +431,16 @@ plot_usmap(data = map_data_combined, regions = "states", values = "winner") +
     legend.text = element_text(size = 10)
   )
 
+election_map
+
+ggsave(
+  filename = "election_map.png",        # File name
+  plot = election_map,                  # Plot object
+  path = "other/plots",               # Directory to save (create if it doesn't exist)
+  width = 10,                            # Width in inches
+  height = 6,                            # Height in inches
+  dpi = 300                              # Resolution
+)
 
 # Chance of victory over time
 

### 2024-11-03 paper/paper.qmd
fixed up results

@@
 
-Our model predicts that Harris has a 63.8% chance of winning the US presidential election.
-While Trump only has a 36.2% chance of winning.
-On average across the 1,000 simulations Harris wins 275.73 Electoral Votes, while Trump wins on average 259.27 votes.
+Our model predicts that Harris has a 68.8% chance of winning the US presidential election.
+While Trump only has a 30.1% chance of winning.
+There is also a 1.1% chance of a tie On average across the 1,000 simulations Harris wins 279.21 Electoral Votes, while Trump wins on average 258.79 votes.
 
 ```{r}
+election_results <- data.frame(
+  Metric = c("Chance of Winning", "Average Electoral Votes"),
+  Harris = c("68.8%", "279.21"),
+  Trump = c("30.1%", "258.79"),
+  stringsAsFactors = FALSE
+)
+
 election_results %>%

### 2024-11-03 scripts/06-bayes_model.R
fixed District of Columbia

@@ -38,6 +38,9 @@ state_stats <- historical_recent %>%
     avg_republican = mean(republican, na.rm = TRUE)
   )
 
+# Replace NA in the 'state' column with "District of Columbia"
+state_stats$state[is.na(state_stats$state)] <- "District of Columbia"
+
 # Separate polling data for Trump and Harris
 poll_data_trump <- poll_data %>%
   filter(candidate == "Trump") %>%
@@ -59,16 +62,12 @@ states_with_poll <- union(states_with_poll_trump, states_with_poll_harris)
 
 # States with no polling data
 states_without_poll <- setdiff(all_states, states_with_poll)
-print(states_without_poll)
 
 #### Merge state_stats with Electoral Votes ####
 
 # Merge state_stats with ev_votes using left_join to retain all states
 state_evs <- ev_votes %>%
   left_join(state_stats, by = "state")
-
-# Verify the merged dataframe
-print(head(state_evs))
 
 #### Define Model Formula ####
 fixed_formula <- support ~ days_to_election + sample_size + transparency_score + pollscore + state
@@ -399,7 +398,7 @@ state_win_counts <- state_win_counts %>%
 
 us_map <- us_map(regions = "states")
 
-#View(us_map)
+View(us_map)
 #View(state_win_counts)
 # Merge map data with state win counts
 map_data_combined <- us_map %>%
@@ -441,215 +440,4 @@ ggsave(
   height = 6,                            # Height in inches
   dpi = 300                              # Resolution
 )
-
-# Chance of victory over time
-
-# Define the maximum number of days before the election
-max_days <- 105
-
-# Define the interval size
-interval_size <- 10
-
-# Create break points for cumulative groups
-break_points <- seq(interval_size, max_days, by = interval_size)
-
-# Ensure the last group includes all remaining days
-if (max_days %% interval_size != 0) {
-  break_points <- c(break_points, max_days)
-}
-
-# Assign each poll to a cumulative group based on days_to_election
-poll_data <- poll_data %>%
-  mutate(
-    group = cut(
-      days_to_election,
-      breaks = c(-Inf, break_points),
-      labels = break_points,
-      right = FALSE
-    )
-  ) %>%
-  filter(!is.na(group)) %>%  # Remove any polls beyond the defined groups
-  mutate(
-    group = as.numeric(as.character(group))
-  )
-
-
-# Function to calculate mean predictors for a candidate within a group
[…]

### 2024-11-03 scripts/05-model_data.R
cleaned up file

@@ -88,18 +88,6 @@ poll_data <- read_csv("data/cleaned_poll_data.csv")
 
 #View(poll_data)
 
-#TODO add lm model, response: pct, predictors: everything else
-
-#Then put the polls through the lm to recover the adjusted average pct for each poll
-
-#Do this for each state and each candidate
-
-#Use this to calculate the number of electoral college votes both candidates
-#will get
-
-# TODO Create a baysian model
-# TODO Make a training/testing split, ideally using CV
-
 ### Model data ####
 harris_models <- list()
 trump_models <- list()

### 2024-11-03 scripts/05-model_data.R
moved electoral votes data frame to the proper file

@@ -85,6 +85,7 @@ region_mapping <- c(
   "Massachusetts" = "NEngland"
 )
 poll_data <- read_csv("data/cleaned_poll_data.csv")
+electoral_votes <- read.csv("data/electoral_votes.csv")
 
 #View(poll_data)
 
@@ -189,38 +190,7 @@ for (state_name in states) {
   
 }
 
-
-#TODO Missing states for some reason, need to check why
 View(state_results)
-
-
-electoral_votes <- data.frame(
-  state = c(
-    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado",
-    "Connecticut", "Delaware", "Florida", "Georgia",
-    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
-    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota",
-    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
-    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
-    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina",
-    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
-    "Washington", "West Virginia", "Wisconsin", "Wyoming"
-  ),
-  ev = c(
-    9, 3, 11, 6, 54, 10,
-    7, 3, 30, 16,
-    4, 4, 19, 11, 6, 6, 8,
-    8, 4, 10, 11, 15, 10,
-    6, 10, 4, 5, 6, 4,
-    14, 5, 28, 16, 3,
-    17, 7, 8, 19, 4, 9,
-    3, 11, 40, 6, 3, 13,
-    12, 4, 10, 3
-  )
-)
-View(electoral_votes)
-
-write_csv(electoral_votes, "data/electoral_votes.csv")
 
 # Merge state_results with electoral votes
 state_results_ev <- merge(state_results, electoral_votes, by = "state")

### 2024-11-03 paper/paper.qmd

@@
 
-FOR IDEALIZED SURVEY
-!TODO like the other appendix, need to talk about depth about the techniques.
-Need to create a google form that is the actual survey.
-Every decision made about the idealized survey must be justified with literature.
-Every detail of the survey must be accounted for.
-Need to write about how the survey was created, and identify and justify all of the tradeoff made.
-TODO!
-
 # Appendix {#sec-appendices}
@@
 In the case of this poll, the number of voters per state region was used to determine the proportion of individuals contacted [@fox].
-This means a greater number of residents in larger states, such as California, would be contacted compared to smaller states, such as Hawaii.
-This method of sampling ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.
+This ensures that a greater number of residents in larger states, such as California, would be contacted compared to smaller states, such as Hawaii.
+This method of sampling allows for a better match between the voices represented by poll results to the ones involved with the real election.
 
@@
 ## Idealized Methodology and Survey
+
+FOR IDEALIZED SURVEY
+!TODO like the other appendix, need to talk about depth about the techniques.
+Need to create a google form that is the actual survey.
+Every decision made about the idealized survey must be justified with literature.
+Every detail of the survey must be accounted for.
+Need to write about how the survey was created, and identify and justify all of the tradeoff made.
+TODO!
+
+Given a budget of $100k and the necessary resources, the idealized methodology for a survey that aims to forecast the results of the US election should be one that targets the states of interest with higher accuracy.
+Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past eight elections [@blueredstate].
+Out of the 51 states, there are 20 states which voted for the same party each time in the past nine elections [@blueredstate].
+On the other hand, there are seven states which have voted for each party three of more times——these would be the states of greater interest to us [@blueredstate].
+Resources can be conserved by focussing our sampling and polling efforts on swing states where there is a greater uncertainty in vote outcome.
+
+After analysis of poll responses, the votes of the consistent states should be added back into consideration before making the final forecast for election results.
+
+A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
 

### 2024-11-04 paper/paper.qmd

@@
 
-=======
-## Analysis into a Pollster's Methodology
+## Analysis into a Pollster's Methodology {#sec-pollster-analysis}
 
@@
 
-Given a budget of $100k and the necessary resources, the idealized methodology for a survey that aims to forecast the results of the US election should be one that targets the states of interest with higher accuracy.
+Given a budget of $100k and the necessary resources, the idealized methodology for a survey that aims to forecast the results of the US election should be one that uses quota sampling and the probability proportionate to size method to target the states of interest specifically.
 Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past eight elections [@blueredstate].
-Out of the 51 states, there are 20 states which voted for the same party each time in the past nine elections [@blueredstate].
-On the other hand, there are seven states which have voted for each party three of more times——these would be the states of greater interest to us [@blueredstate].
-Resources can be conserved by focussing our sampling and polling efforts on swing states where there is a greater uncertainty in vote outcome.
+Out of the 51 states, there are 20 states which voted for the same party each time in the past nine elections and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
+On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections——these would be the states of greater interest to us [@blueredstate].
+
+Using the non-probability sampling method of quota sampling, these swing states will first be selected.
+This targetted sampling process allows us to focus our polling efforts on gathering more data from the areas where there is higher uncertainty, but the tradeoff is bias is easily introduced with the assumptions we are making, potentially impacting the accuracy of our forecast [@quotasampling].
+This means that more work is required to adjust for these biases while also minimizing the inflation of variance [@quotasamping].
+
+Next, within these states, probability proportionate to size will be used to select participants from the nationwide file of registered voters to ensure that the number of people polled is proportional to the number of voters per region.
+This is the sampling method used in @fox, as discussed above in @sec-pollster-analysis, and it allows for greater accuracy of prediction due to each unit's strategically adjusted probability of inclusion [@Latpate2021].
+
+- questions (wording, order, options)
+- treatment of non-response
+- offer of incentive
 

### 2024-11-04 paper/paper.qmd
Update abstract and subtitle

@@
 title: "Forecasting the United States 2024 Presidential Election"
-subtitle: "Kamala Harris wins 60th presidency with 265 Electoral College Votes"
+subtitle: "Kamala Harris wins 60th presidency with 279 Electoral College Votes"
 author: 
@@
 From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system.
-The models forecast Kamala Harris as the winner of election, winning 265 electoral college votes versus the 172 votes Donald Trump acquires.
+The models forecast Kamala Harris as the winner of election, winning 279 electoral college votes versus the 258 votes Donald Trump acquires.
 Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process.
@@
 
--Diagnostics
+## Diagnostics {#sec-diagnostics}
 

### 2024-11-04 paper/paper.qmd
added swing state chart

--- 2024-11-03 paper/paper.qmd
@@
            show_col_types = FALSE)
+
+swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))
 ```
@@
 
-- **support<sub>i</sub>**  
+- **Support**  
   *Definition:* The support level for candidate \( i \) in a given poll.  
@@
 
-- **sample\_size<sub>i</sub>**  
+- **Sample Size**  
   *Definition:* Represents the number of respondents in poll \( i \).  
@@
 
-- **days\_to\_election<sub>i</sub>**  
+- **Days to election**  
   *Definition:* Denotes the number of days remaining until the election when poll \( i \) was conducted.  
@@
 
-- **transparency\_score<sub>i</sub>**  
+- **Transparency Score**  
   *Definition:* A numerical score reflecting the pollster’s transparency regarding their methodology, ranging up to 10, with higher scores indicating greater transparency.  
@@
 
-- **pollscore<sub>i</sub>**  
+- **Pollscore**  
   *Definition:* A numeric value representing the reliability and bias of the pollster, where negative values denote better reliability.  
@@
 
-- **state<sub>i</sub>**  
+- **State**  
   *Definition:* Captures state-specific effects as fixed effects.  
   *Justification:* Allows the model to account for regional variations in support that are not explained by other predictors, incorporating state-specific nuances.
-
-- **ϵ<sub>i</sub>**  
-  *Definition:* The error term, assumed to follow a normal distribution.  
-  *Justification:* Represents unexplained variability in the model, ensuring that the residuals meet the assumptions of the statistical analysis.
-
-Variable Definitions:
-
-- **Sample Size (\(\text{sample\_size}\))**: The size of the poll is included as larger samples generally provide more accurate estimates of support, reducing sampling variability.
-
-- **Days to Election (\(\text{days\_to\_election}\))**: This variable accounts for the temporal proximity to the election, as support levels may fluctuate as the election approaches, capturing trends in voter sentiment over time.
-
-- **Transparency Score (\(\text{transparency\_score}\))**: A higher transparency score indicates greater disclosure of poll methodology, potentially correlating with the reliability and credibility of the poll results.
-
-- **Pollscore (\(\text{pollscore}\))**: This score represents the inherent bias and error of the pollster, with lower (more negative) values indicating higher reliability.
-It accounts for systematic deviations in poll results based on historical pollster performance [REF 538 METHOD].
-
-- **State (\(\text{state}\))**: Including state as a fixed effect allows the model to account for regional variations in support that are not captured by other predictors, ensuring state-specific nuances are incorporated into the support estimates.
 
@@
 Alternatively, imputation methods or political regional averages could have been used, but historical averages are used for their simplicity and relevance.
+
+Though Maine and Nebraska allow for their Electoral College votes to be split.
+In Maine two of its four votes are awarded statewide and the other two for each of its congressional districts.
+In Nebraska two of its five votes are awarded statewide and the other three for each of its congressional districts.
+Our model simplifies this dynamic and all of their votes are assigned according the support overall in the state.
+The main reason for this is that some of the districts have little or no polling data, making it difficult to measure their support for either candidate.
 
@@
 knitr::include_graphics(plot_path)
+
+```
+
+The results of the US elections in large part hinge on the direction of the swing states.
+A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris.
+[FIG] shows the probability that both candidates have of winning each of the 7 swing states.
+
+```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
+# Reshape the data for plotting
+swing_long <- swing_probs %>%
+  pivot_longer(cols = c("Trump", "Harris"), names_to = "Candidate", values_to = "Probability")
+
+custom_colors <- c("Trump" = "#DD1717", "Harris" = "#0F4392")
+
+# Create the bar plot
+ggplot(swing_long, aes(x = state, y = Probability, fill = Candidate)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_fill_manual(values = custom_colors) + 
+  labs(
+    title = "Probability of Winning Swing States",
+    x = "Swing State",
+    y = "Probability (%)",
+    fill = "Candidate"
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 12),
+    plot.title = element_text(hjust = 0.5)
+  )
 

### 2024-11-04 scripts/03-clean_data.R

@@ -197,6 +197,8 @@ missing <- poll_data %>%
     missing_state = sum(is.na(state))
   )
 
+cleaned_poll_data <- cleaned_poll_data |> filter(numeric_grade >= 1.5)
+
 #View(missing)
 
 #View(cleaned_historical_state_data)

### 2024-11-04 paper/paper.qmd
fixed

@@
 ---
-![State Election Predictions](other/plots/election_map.png){width=80%}
 

### 2024-11-04 scripts/06-bayes_model.R

@@ -274,6 +274,59 @@ avg_harris_ev <- mean(harris_ev_counts)
 cat("Average Electoral Votes for Trump:", round(avg_trump_ev, 2), "\n")
 cat("Average Electoral Votes for Harris:", round(avg_harris_ev, 2), "\n")
 
+
+# Define the seven swing states
+swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
+
+# Initialize a data frame to store the probabilities
+swing_probabilities <- data.frame(
+  state = swing_states,
+  Trump = NA_real_,
+  Harris = NA_real_,
+  stringsAsFactors = FALSE
+)
+
+# Loop through each swing state to calculate win probabilities
+for (state in swing_states) {
+  if (state %in% all_states) {
+    # Retrieve the support vectors for the current state across all simulations
+    trump_support <- trump_support_matrix[, state]
+    harris_support <- harris_support_matrix[, state]
+    
+    # Calculate the number of simulations where Trump wins the state
+    trump_wins <- sum(trump_support > harris_support, na.rm = TRUE)
+    
+    # Calculate the number of simulations where Harris wins the state
+    harris_wins <- sum(harris_support > trump_support, na.rm = TRUE)
+    
+    # Calculate the probabilities as percentages
+    trump_prob <- (trump_wins / num_simulations) * 100
+    harris_prob <- (harris_wins / num_simulations) * 100
+    
+    # Assign the probabilities to the data frame
+    swing_probabilities[swing_probabilities$state == state, "Trump"] <- trump_prob
+    swing_probabilities[swing_probabilities$state == state, "Harris"] <- harris_prob
+  } else {
+    # If the state is not present in the data, assign NA
+    swing_probabilities[swing_probabilities$state == state, "Trump"] <- NA
+    swing_probabilities[swing_probabilities$state == state, "Harris"] <- NA
+    warning(paste("State", state, "is not present in the dataset."))
+  }
+}
+
+# Optionally, format the probabilities to two decimal places
+swing_probabilities <- swing_probabilities %>%
+  mutate(
+    Trump = round(Trump, 2),
+    Harris = round(Harris, 2)
+  )
+
+# Print the swing states probability table
+print(swing_probabilities)
+
+# Optionally, save the table to a CSV file for further analysis or reporting
[…]

### 2024-11-04 paper/paper.qmd
Update paper.qmd

@@
 Additionally, it helps predict potential policy shifts and holds political institutions accountable, contributing to a more transparent and informed democracy."
-format: pdf
+format:
+  pdf:
+    prefer-html: true
+  html: default
 header-includes:
@@
 library(ggplot2)
+library(webshot2)
+library(knitr)
+opts_knit$set(webshot = TRUE)
 
@@
            show_col_types = FALSE)
-
-swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))
 ```
@@
 
-- **Support**  
+- **support<sub>i</sub>**  
   *Definition:* The support level for candidate \( i \) in a given poll.  
@@
 
-- **Sample Size**  
+- **sample\_size<sub>i</sub>**  
   *Definition:* Represents the number of respondents in poll \( i \).  
@@
 
-- **Days to election**  
+- **days\_to\_election<sub>i</sub>**  
   *Definition:* Denotes the number of days remaining until the election when poll \( i \) was conducted.  
@@
 
-- **Transparency Score**  
+- **transparency\_score<sub>i</sub>**  
   *Definition:* A numerical score reflecting the pollster’s transparency regarding their methodology, ranging up to 10, with higher scores indicating greater transparency.  
@@
 
-- **Pollscore**  
+- **pollscore<sub>i</sub>**  
   *Definition:* A numeric value representing the reliability and bias of the pollster, where negative values denote better reliability.  
@@
 
-- **State**  
+- **state<sub>i</sub>**  
   *Definition:* Captures state-specific effects as fixed effects.  
   *Justification:* Allows the model to account for regional variations in support that are not explained by other predictors, incorporating state-specific nuances.
+
+- **ϵ<sub>i</sub>**  
+  *Definition:* The error term, assumed to follow a normal distribution.  
+  *Justification:* Represents unexplained variability in the model, ensuring that the residuals meet the assumptions of the statistical analysis.
+
+Variable Definitions:
+
+- **Sample Size (\(\text{sample\_size}\))**: The size of the poll is included as larger samples generally provide more accurate estimates of support, reducing sampling variability.
+
+- **Days to Election (\(\text{days\_to\_election}\))**: This variable accounts for the temporal proximity to the election, as support levels may fluctuate as the election approaches, capturing trends in voter sentiment over time.
+
+- **Transparency Score (\(\text{transparency\_score}\))**: A higher transparency score indicates greater disclosure of poll methodology, potentially correlating with the reliability and credibility of the poll results.
+
+- **Pollscore (\(\text{pollscore}\))**: This score represents the inherent bias and error of the pollster, with lower (more negative) values indicating higher reliability.
+It accounts for systematic deviations in poll results based on historical pollster performance [REF 538 METHOD].
+
+- **State (\(\text{state}\))**: Including state as a fixed effect allows the model to account for regional variations in support that are not captured by other predictors, ensuring state-specific nuances are incorporated into the support estimates.
 
@@
 Alternatively, imputation methods or political regional averages could have been used, but historical averages are used for their simplicity and relevance.
-
-Though Maine and Nebraska allow for their Electoral College votes to be split.
-In Maine two of its four votes are awarded statewide and the other two for each of its congressional districts.
-In Nebraska two of its five votes are awarded statewide and the other three for each of its congressional districts.
-Our model simplifies this dynamic and all of their votes are assigned according the support overall in the state.
-The main reason for this is that some of the districts have little or no polling data, making it difficult to measure their support for either candidate.
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 election_results <- data.frame(
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 knitr::include_graphics(plot_path)
-
-```
-
-The results of the US elections in large part hinge on the direction of the swing states.
-A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris.
-[FIG] shows the probability that both candidates have of winning each of the 7 swing states.
-
-```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
-# Reshape the data for plotting
-swing_long <- swing_probs %>%
-  pivot_longer(cols = c("Trump", "Harris"), names_to = "Candidate", values_to = "Probability")
-
-custom_colors <- c("Trump" = "#DD1717", "Harris" = "#0F4392")
-
-# Create the bar plot
-ggplot(swing_long, aes(x = state, y = Probability, fill = Candidate)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  scale_fill_manual(values = custom_colors) + 
-  labs(
-    title = "Probability of Winning Swing States",
-    x = "Swing State",
-    y = "Probability (%)",
-    fill = "Candidate"
-  ) +
-  theme_minimal() +
-  theme(
-    text = element_text(size = 12),
-    plot.title = element_text(hjust = 0.5)
-  )
 
@@
 
-!TODO need to add details about the techniques used by the pollster.
-For example, if stratified sampling was used, dive into what is stratified sampling.
-What is it good for why is works.
-Reference a academic paper about stratified sampling.
-TODO!
-
 ## Analysis into a Pollster's Methodology {#sec-pollster-analysis}
@@
 ## Idealized Methodology and Survey
-
-FOR IDEALIZED SURVEY
-!TODO like the other appendix, need to talk about depth about the techniques.
-Need to create a google form that is the actual survey.
-Every decision made about the idealized survey must be justified with literature.
-Every detail of the survey must be accounted for.
-Need to write about how the survey was created, and identify and justify all of the tradeoff made.
-TODO!
 
@@
 ```
+\newpage
+
 # References

### 2024-11-04 paper/paper.qmd
test
added barchart

--- 2024-11-04 paper/paper.qmd
@@
     title = "Probability of Winning Swing States",
-    x = "Swing State",
+    x = "State",
     y = "Probability (%)",

### 2024-11-04 paper/paper.qmd
Update paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 ```
+The results of the US elections in large part hinge on the direction of the swing states.
+A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris.
+[FIG] shows the probability that both candidates have of winning each of the 7 swing states.
+
+```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
+# Reshape the data for plotting
+swing_long <- swing_probs %>%
+  pivot_longer(cols = c("Trump", "Harris"), names_to = "Candidate", values_to = "Probability")
+
+custom_colors <- c("Trump" = "#DD1717", "Harris" = "#0F4392")
+
+# Create the bar plot
+ggplot(swing_long, aes(x = state, y = Probability, fill = Candidate)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_fill_manual(values = custom_colors) +
+  labs(
+    title = "Probability of Winning Swing States",
+    x = "State",
+    y = "Probability (%)",
+    fill = "Candidate"
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 12),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 

### 2024-11-04 paper/paper.qmd
fixed

--- 2024-11-04 paper/paper.qmd
@@
 
-plot_path <- here::here("other", "plots", "election_map.png")
+plot_path <- here::here("other", "plots", "election_map.pdf")
 
@@
 election_results %>%
-  kable("html", caption = "Election Prediction Results") %>%
+  kable(caption = "Election Prediction Results") %>%
   kable_styling(

### 2024-11-04 paper/paper.qmd
Update paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 
-Given a budget of $100k and the necessary resources, the idealized methodology for a survey that aims to forecast the results of the US election should be one that uses quota sampling and the probability proportionate to size method to target the states of interest specifically.
+In an ideal world, where a $100k in funds and all necessary resources could be obtained from a neutral source for the purpose of polling the population, the idealized methodology for a survey that aims to forecast the results of the US election should be one that uses quota sampling and the probability proportionate to size method to target the states of interest specifically.
 Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past eight elections [@blueredstate].
@@
 
+After the sample has been selected with the above methods, the poll itself will be conducted through a survey sent to the chosen participants in order to reduce interviewer bias, or the effect of the interviewer’s own views on the measured responses, which can affect results when face-to-face or phone interviews are utilized [@citeTB].
+As well, to reduce social desirability effect, the survey must emphasize in the beginning that the respondents’ identity will be kept anonymous to the researchers and the final reported data [@citeStantcheva].
+
 - questions (wording, order, options)
@@
 A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
+In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, the question order should be randomized when being presented to different respondents.
+This will help reduce response order bias, which is when the answer of a question is affected by the order in which it appeared in the survey [@citeStantcheva].
 

### 2024-11-04 paper/paper.qmd
added echo false

--- 2024-11-04 paper/paper.qmd
@@
 ```{r}
+#| fig-cap: Election Prediction Results
+#| echo: false
+#| warning: false
+#| message: false
+
 election_results <- data.frame(
@@
 ```{r}
+#| fig-cap: Election Map
+#| echo: false
+#| warning: false
+#| message: false
 knitr::include_graphics(plot_path)
@@
 ```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
+#| echo: false
+#| warning: false
+#| message: false
 # Reshape the data for plotting

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
            show_col_types = FALSE)
+swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))
 ```
@@
 ```{r}
+#| fig-cap: Election Prediction Results
+#| echo: false
+#| warning: false
+#| message: false
+
 #| echo: false
@@
 ```{r}
+#| fig-cap: Election Map
 #| echo: false

### 2024-11-04 paper/paper.qmd

@@
 
-Based on the simulated elections, this map illustrates the most likely winner of each state. In this scenario Harris wins 277 Electoral Votes, and Trump wins 262.[Appendix]
+Based on the simulated elections, this map illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.[Appendix]
 
@@
 ## Betting Markets {#sec-disc-bet}
+
+Prediction betting markets provide a new data source that may be valuable for predicting the outcome of the US election.
+These online betting markets allow people to buy shares that represent an outcome of some event.
+If the event occurs then the share is worth $1, if not then the share's price falls to 0.
+This means the live share price reflects the probability of the event occurring.
+For example, for the outcome of the presidential election market, if Trump's share price is 65 cents this means the market is predicting there is a 65% chance that Trump will win the election.
+Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
+Some papers have included data from these markets in their models to improve their prediction accuracy.[REF https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+
+Advantages
+
+Disadvantages and drawbacks
 

### 2024-11-04 scripts/03-clean_data.R
updated data

@@ -187,16 +187,6 @@ print(missing_states)
 #!CHECKING STATES!#
 
 
-# Summarize missing values for each predictor
-missing <- poll_data %>%
-  summarize(
-    missing_days_to_election = sum(is.na(days_to_election)),
-    missing_sample_size = sum(is.na(sample_size)),
-    missing_transparency_score = sum(is.na(transparency_score)),
-    missing_pollscore = sum(is.na(pollscore)),
-    missing_state = sum(is.na(state))
-  )
-
 cleaned_poll_data <- cleaned_poll_data |> filter(numeric_grade >= 1.5)
 
 #View(missing)

### 2024-11-04 paper/paper.qmd
added survey monkey question randomization feature

@@
 A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
-In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, the question order should be randomized when being presented to different respondents.
+In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, such as Survey Monkey, the question order would be randomized when being presented to different respondents with the Question Randomization feature [@surveymonkey].
 This will help reduce response order bias, which is when the answer of a question is affected by the order in which it appeared in the survey [@citeStantcheva].

### 2024-11-04 scripts/03-clean_data.R
run model

@@ -187,7 +187,7 @@ print(missing_states)
 #!CHECKING STATES!#
 
 
-cleaned_poll_data <- cleaned_poll_data |> filter(numeric_grade >= 1.5)
+
 
 #View(missing)
 

### 2024-11-04 paper/paper.qmd
added details

--- 2024-11-04 paper/paper.qmd
@@
 For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
+This data was collected from a github repository [REF] who collected their data from Wikipedia.
 This decision is based on the assumption that certain states are very unlikely to change their voting patterns, reducing the necessity for current polling data.
@@
 We evaluated model convergence using the R-hat statistic and trace plots.
-All R-hat values for model parameters were below 1.1, indicating satisfactory convergence.
 Trace plots revealed well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process [REF].
@@
 
-Residual plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines.
+Residual plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines[REF].
 This pattern may indicate unmodeled heterogeneity or data limitations, suggesting areas where the model's fit could be improved.
@@
 
-Our model predicts that Harris has a 68.8% chance of winning the US presidential election.
-While Trump only has a 30.1% chance of winning.
-There is also a 1.1% chance of a tie On average across the 1,000 simulations Harris wins 279.21 Electoral Votes, while Trump wins on average 258.79 votes.
+Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
+While Trump only has a 31.7% chance of winning.
+There is also a 0.9% chance of a tie On average across the 1,000 simulations Harris wins 279.33 Electoral Votes, while Trump wins on average 258.67 votes.
 
@@
   Metric = c("Chance of Winning", "Average Electoral Votes"),
-  Harris = c("68.8%", "279.21"),
-  Trump = c("30.1%", "258.79"),
+  Harris = c("67.4%", "279.33"),
+  Trump = c("31.7%", "258.67"),
   stringsAsFactors = FALSE
@@
 
+Our model does not take into account state correlations based on past election data and research based insights into voter preference.
+For example a model developed by Chernov, Elenev, and Song, posits if Harris were to win Pennsylvania, then her chances of winning Nevada would increase from 24% to 68% [https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+These correlations between states could have a large impact on election predictions and our model is limited by not modelling these interactions.
+
 ## Next Steps {#sec-disc-next}
@@
 Additionally, enlisting high-profile figures or celebrities, like Elon Musk or LeBron James—who have recently endorsed Trump and Harris, respectively—could enhance their appeal in these swing states.
-These endorsements would help candidates reach diverse and niche demographics, broadening their voter base and potentially swaying undecided voters.
+These endorsements would help candidates reach diverse and niche demographics, broadening teheir voter base and potentially swaying undecided voters.
 
@@
 
-Advantages
+Some key advantages of these markets is that they are real time, so the market adjusts quickly to important and material news.
+Where as poll are sluggish to incorporate big changes as new polls need to be sent out and analysed.
+Another advantage is that these markets may attract sophisticated investors that could employ methods that differ from polling but still offer predictive capability.
+The larger the trade volume the more likely these actors are to be present and wash out any unsophisticated trading.
 
-Disadvantages and drawbacks
+On the other hand the main disadvantage is that all of the investors in this market essentially completely anonymous.
+This means there is no way of checking whether a large investor, who has significant price setting power, is indeed sophisticated or not.
+Furthermore, because these markets have only recently become popular or even legal in some cases one might expect these markets to be full of amateurs.
+These markets may display a bais as well, where the largest market Polymarket [REF] is based on crypto-currency, and Trump is popular among crypto-enthusiasts, and large financial institutions are underrepresented in the crypto-market.
+
+This is a new frontier in forecasting and including data from these markets may prove beneficial. Or these markets may be nothing more than a hobby, and hold no useful data.
 

### 2024-11-04 paper/paper.qmd
finished section on ideal poll questions design

--- 2024-11-04 paper/paper.qmd
@@
 
-- questions (wording, order, options)
-- treatment of non-response
-- offer of incentive
+The wording of questions should be straightforward and easy to understand, minimizing error in measurement caused by misinterpretation of questions.
+The question should also be asked neutrally, without leading participants to any particular response.
+Likert scales, including an option for “don’t know”, should be used for all questions for ease and consistency of participant understanding.
+As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB].
+The “don’t know” option is also effective in helping reduce the decision difficulty to certain participants who are truly unsure.
+Before officially deploying the survey with our selected sample, it first needs to be tested to catch ambiguous or other poorly phrased questions [@citeTB].
+To encourage participation in this research, a financial incentive should be offered for full completion of the poll.
 
-After analysis of poll responses, the votes of the consistent states should be added back into consideration before making the final forecast for election results.
+After analysis of poll responses, the votes of the historically consistent states should be added back into consideration before making the final forecast for election results.
 

### 2024-11-04 paper/paper.qmd
added voter file inclusions

@@
   - Michelle Ji
-  - Cher Ning
+  - Cher Ning-Li
 thanks: "Code and data are available at: https://github.com/Ford-Robert/us-presidential-election."
@@
 Out of the 51 states, there are 20 states which voted for the same party each time in the past nine elections and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
-On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections——these would be the states of greater interest to us [@blueredstate].
+On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections—these would be the states of greater interest to us [@blueredstate].
 
@@
 
+Some information of interest to us, such as voting history and racial identification, is already included in the registered voter file, so these questions can be excluded from the poll itself [@voterfiledata].
 After analysis of poll responses, the votes of the historically consistent states should be added back into consideration before making the final forecast for election results.

### 2024-11-04 paper/paper.qmd
updated paper to include appendix graphs

--- 2024-11-04 paper/paper.qmd
@@
 
-plot_path <- here::here("other", "plots", "election_map.pdf")
-
 #read in cleaned data
@@
 swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))
+
+state_win_counts <- read_csv(file = here("other/plots/state_win_counts.csv"))
 ```
@@
 
-Posterior predictive checks were conducted to assess the model's ability to replicate observed data.
+Posterior predictive checks were conducted to assess the model's ability to replicate observed data[REF].
 Density overlay plots for both Trump and Harris models indicated reasonable fits, although some discrepancies were noted near the peaks of the distributions.
 Specifically, the Trump model's replicated data tended to be slightly smaller around the peak, while the Harris model exhibited more extreme differences at the peak.
-Residual plots appeared randomly scattered, suggesting no major systematic biases, although some clumping was observed, indicating areas where the model may fit less effectively [REF].
 
@@
 ```{r}
+#| label: fig-4
 #| fig-cap: Election Prediction Results
@@
 
-#| echo: false
-#| warning: false
-#| message: false
 election_results <- data.frame(
@@
 
-# Display a nicely formatted table
-kable(election_results, caption = "Election Prediction Results")
-
 ```
 
-Based on the simulated elections, this map illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.[Appendix]
+Based on the simulated elections, @fig-5 illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.
 
 ```{r}
+#| label: fig-5
 #| fig-cap: Election Map
@@
 #| message: false
+
+plot_path <- here::here("other", "plots","election_map.pdf" )
 knitr::include_graphics(plot_path)
@@
 A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris.
-[FIG] shows the probability that both candidates have of winning each of the 7 swing states.
+@fig-6 shows the probability that both candidates have of winning each of the 7 swing states.
 
 ```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
+#| label: fig-6
 #| echo: false
@@
 
-The wording of questions should be straightforward and easy to understand, minimizing error in measurement caused by misinterpretation of questions.
-The question should also be asked neutrally, without leading participants to any particular response.
-Likert scales, including an option for “don’t know”, should be used for all questions for ease and consistency of participant understanding.
-As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB].
-The “don’t know” option is also effective in helping reduce the decision difficulty to certain participants who are truly unsure.
-Before officially deploying the survey with our selected sample, it first needs to be tested to catch ambiguous or other poorly phrased questions [@citeTB].
-To encourage participation in this research, a financial incentive should be offered for full completion of the poll.
+- questions (wording, order, options)
+- treatment of non-response
+- offer of incentive
 
-After analysis of poll responses, the votes of the historically consistent states should be added back into consideration before making the final forecast for election results.
+After analysis of poll responses, the votes of the consistent states should be added back into consideration before making the final forecast for election results.
 
@@
 ```
+
+## State By State Win {#sec-trace-plots}
+
+```{r}
+#| fig-cap: Number of Wins in 1000 Simulations by State
+#| echo: false
+#| warning: false
+#| message: false
+
+kable(state_win_counts, caption = "Election Results by State", align = 'lccc') %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
+
+```
+
+## Trace Plots {#sec-trace-plots}
+
+```{r}
+#| fig-cap: Trump Trace Plots
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","trump_trace.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
+```{r}
+#| fig-cap: Harris Trace Plots
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","harris_trace.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
+## Posterior Predictive Checks {#sec-ppc}
+
+```{r}
+#| fig-cap: Trump Posterior Predictive Checks
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","trump_ppc.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
+```{r}
+#| fig-cap: Harris Posterior Predictive Checks
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","harris_ppc.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
+## Residual Plots {#sec-residual-plots}
+
+```{r}
+#| fig-cap: Trump Residuals
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","trump_residuals.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
+```{r}
+#| fig-cap: Harris Residuals
+#| echo: false
+#| warning: false
+#| message: false
+
+plot_path <- here::here("other", "plots","harris_residuals.pdf" )
+knitr::include_graphics(plot_path)
+
+```
+
 \newpage

### 2024-11-04 paper/paper.qmd
updated ideal poll

--- 2024-11-04 paper/paper.qmd
@@
 The question should also be asked neutrally, without leading participants to any particular response.
-Likert scales, including an option for “don’t know”, should be used for all questions for ease and consistency of participant understanding.
+Likert scales, including an option for “don’t know”, should be used whenever possible for ease and consistency of participant understanding.
 As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB].
@@
 To encourage participation in this research, a financial incentive should be offered for full completion of the poll.
+Displaying the progress bar and indicating the survey length at the very beginning also helps encourage participants to complete the full questionnaire.
+Our poll would also be submitted to the appropriate ethics review boards to increase confidence in our methodology.
 
@@
 A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
-In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, such as Survey Monkey, the question order would be randomized when being presented to different respondents with the Question Randomization feature [@surveymonkey].
+In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, the question order would be randomized when being presented to different respondents, with consideration for related questions must be kept together.
 This will help reduce response order bias, which is when the answer of a question is affected by the order in which it appeared in the survey [@citeStantcheva].

### 2024-11-04 paper/paper.qmd
minor formatting + citation edits
bib refrences + tabel of no poll data states

--- 2024-11-04 paper/paper.qmd
@@
 Ideally we would fit both models and use diagnostics to choose which has lower bais and variance, and thus better predictive quality.
-Additionally, simpler models excluding variables like transparency_score and pollscore were evaluated but found inadequate in capturing the state-by-state differences in the transparency and reliability of the polls, potentially leading to biased support estimates.
+Additionally, simpler models excluding variables like Transparency Score and Poll Score were evaluated but found inadequate in capturing the state-by-state differences in the transparency and reliability of the polls, potentially leading to biased support estimates.
 
@@
 For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
-This data was collected from a github repository [REF] who collected their data from Wikipedia.
+This data was collected from a github repository (@hist_state) who collected their data from Wikipedia.
 This decision is based on the assumption that certain states are very unlikely to change their voting patterns, reducing the necessity for current polling data.
-The ten states that do not have polling data [REF] have not changed parties in at least the last 20 plus years.
-This may be why they do not have polling data, as it would be a waste of resources for agencies to poll them.
+The ten states that do not have polling data have not changed parties in at least the last 20 plus years (@sec-voting-hist).
 Alternatively, imputation methods or political regional averages could have been used, but historical averages are used for their simplicity and relevance.
@@
 The Bayesian models were implemented using the `rstanarm` package in R, which interfaces with Stan for efficient Markov Chain Monte Carlo (MCMC) sampling.
-Data manipulation and visualization were conducted using the `tidyverse` suite of packages, while model diagnostics leveraged `bayesplot` and other related packages.
+Data manipulation and visualization were conducted using the `tidyverse` suite of packages, while model diagnostics used `bayesplot` and other related packages.
 This combination of tools facilitated a streamlined workflow for model fitting, prediction, and validation.
@@
 We evaluated model convergence using the R-hat statistic and trace plots.
-Trace plots revealed well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process [REF].
+Trace plots, @sec-trace-plots revealed well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process.
 
-Posterior predictive checks were conducted to assess the model's ability to replicate observed data[REF].
+Posterior predictive checks were conducted to assess the model's ability to replicate observed data, @sec-ppc.
 Density overlay plots for both Trump and Harris models indicated reasonable fits, although some discrepancies were noted near the peaks of the distributions.
@@
 
-Residual plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines[REF].
+Residual plots (found in @sec-residual-plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines.
 This pattern may indicate unmodeled heterogeneity or data limitations, suggesting areas where the model's fit could be improved.
@@
 Ideally several models would be created then using model selection techniques would be used to determine the best model for predicting the outcome of the US election.
-Due to time constraints the folling models were considered but not implemented.
-A set of Bayesian models with state as a random effect and models excluding specific predictors like transparency_score.
+Due to time constraints the following models were considered but not implemented.
+A set of Bayesian models with state as a random effect and models excluding specific predictors like Transparency Score.
 More complex models with additional interaction terms or non-linear components were also considered
@@
 
-All analyses were conducted using R (version 4.2.1) with the following key packages:
+All analyses were conducted using R (@citeR) with the following key packages:
 
-- **tidyverse:** For data manipulation and visualization.
-- **rstanarm:** For Bayesian model fitting using Stan.
-- **brms:** Alternative package for Bayesian regression modeling.
-- **bayesplot:** For comprehensive model diagnostics and visualization.
+- **tidyverse:** For data manipulation and visualization. (@tidyverse)
+- **rstanarm:** For Bayesian model fitting using Stan. (@rstanarm1, @rstanarm2)
+- **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
+- **bayesplot:** For comprehensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
 
@@
 
+@table-2 displays each candidates chance of winning and average number of Elector Votes they received over the 1000 simulated elections.
 Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
@@
 ```{r}
-#| label: fig-4
+#| label: table-2
 #| fig-cap: Election Prediction Results
@@
 
-Based on the simulated elections, @fig-5 illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.
+Based on the simulated elections, @fig-5 illustrates the most likely winner of each state.
+In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.
+A table in @sec-state-wins, shows how this map was constructed by taking the most probable winners of each state.
 
@@
 
-## State By State Win {#sec-trace-plots}
+## Deep Red/Blue State Voting History {#sec-voting-hist}
+
+```{r}
+#| fig-cap: Voting History of the 10 States That do not Have Poll Data
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create a data frame with the provided data
+df <- data.frame(
+  State = c("Alabama", "Delaware", "Hawaii", "Idaho", "Illinois",
+            "Kentucky", "Louisiana", "Mississippi", "Tennessee", "Wyoming"),
+  Party = c("Rep", "Dem", "Dem", "Rep", "Dem",
+            "Rep", "Rep", "Rep", "Rep", "Rep"),
+  Start_Year = c(1980, 1992, 1998, 1976, 1992,
+                 2000, 2000, 1980, 2000, 1976),
+  stringsAsFactors = FALSE
+)
+
+# -----------------------------
+# Step 3: Reshape the Data for a Compact Table
+# -----------------------------
+
+# Pair the states into two columns per row
+# We'll split the data into two halves
+df1 <- df[1:5, ]  # First 5 states
+df2 <- df[6:10, ] # Next 5 states
+
+# Combine into a single data frame with 6 columns
+compact_df <- data.frame(
+  State_1 = df1$State,
+  Year_1 = df1$Start_Year,
+  Party_1 = df1$Party,
+  State_2 = df2$State,
+  Year_2 = df2$Start_Year,
+  Party_2 = df2$Party,
+  stringsAsFactors = FALSE
+)
+
+# -----------------------------
+# Step 4: Apply Cell-wise Coloring Based on Party
+# -----------------------------
+
+# Function to assign colors based on party
+assign_color <- function(party) {
+  ifelse(party == "Rep", "#DD1717", "#0F4392")
+}
+
+# Apply cell_spec to each cell that needs coloring
+compact_df <- compact_df %>%
+  mutate(
+    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white"),
+    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white"),
+    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white"),
+    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white")
+  )
+
+# -----------------------------
+# Step 5: Create the Styled Compact Table with Note
+# -----------------------------
+
+# Create and style the table
+styled_table <- compact_df %>%
+  select(State_1, Year_1, State_2, Year_2) %>%
+  kable("html", escape = FALSE, align = "c",
+        col.names = c("State", "Last Time Voted for a Different Party",
+                      "State", "Last Time Voted for a Different Party")) %>%
+  kable_styling("striped", full_width = FALSE, position = "left") %>%
+  # Header styling
+  row_spec(0, bold = TRUE, background = "#D3D3D3") %>%
+  # Add a footnote below the table
+  add_footnote("Note: Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.",
+  )
+
+# Display the table
+styled_table
+```
+## State By State Win {#sec-state-wins}
 

### 2024-11-04 other/deep_states.R

@@ -1,219 +1,0 @@
-# -----------------------------
-# R Script to Create a Compact Styled Table with 4 Columns and a Note
-# -----------------------------
-
-# -----------------------------
-# Step 1: Install and Load Required Packages
-# -----------------------------
-
-# Uncomment the following lines to install the packages if they are not already installed
-# install.packages("kableExtra")
-# install.packages("dplyr")
-
-# Load the necessary libraries
-library(kableExtra)
-library(dplyr)
-
-# -----------------------------
-# Step 2: Define the Data
-# -----------------------------
-
-# Create a data frame with the provided data
-df <- data.frame(
-  State = c("Alabama", "Delaware", "Hawaii", "Idaho", "Illinois",
-            "Kentucky", "Louisiana", "Mississippi", "Tennessee", "Wyoming"),
-  Party = c("Rep", "Dem", "Dem", "Rep", "Dem",
-            "Rep", "Rep", "Rep", "Rep", "Rep"),
-  Start_Year = c(1980, 1992, 1998, 1976, 1992,
-                 2000, 2000, 1980, 2000, 1976),
-  stringsAsFactors = FALSE
-)
-
-# -----------------------------
-# Step 3: Reshape the Data for a Compact Table
-# -----------------------------
-
-# Pair the states into two columns per row
-# We'll split the data into two halves
-df1 <- df[1:5, ]  # First 5 states
-df2 <- df[6:10, ] # Next 5 states
-
-# Combine into a single data frame with 6 columns
-compact_df <- data.frame(
-  State_1 = df1$State,
-  Year_1 = df1$Start_Year,
-  Party_1 = df1$Party,
-  State_2 = df2$State,
-  Year_2 = df2$Start_Year,
-  Party_2 = df2$Party,
-  stringsAsFactors = FALSE
-)
-
-# -----------------------------
-# Step 4: Apply Cell-wise Coloring Based on Party
-# -----------------------------
-
-# Function to assign colors based on party
-assign_color <- function(party) {
-  ifelse(party == "Rep", "red", "blue")
-}
-
-# Apply cell_spec to each cell that needs coloring
-compact_df <- compact_df %>%
-  mutate(
-    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white"),
-    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white"),
-    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white"),
[…]

### 2024-11-04 paper/paper.qmd
fixed up

@@
 
+All model construction and analysis was conducted using R (@citeR) with the following key packages:
+
+- **tidyverse:** For data manipulation and visualization. (@tidyverse)
+- **rstanarm:** For Bayesian model fitting using Stan. (@rstanarm1, @rstanarm2)
+- **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
+- **bayesplot:** For comprehensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
+
 To predict state level support and calculate the expected Electoral Votes (EV) for each candidate in the upcoming US election, we employed a Bayesian linear regression model.
-The model is mathematically specified as follows:
+The model is specified as follows:
 
@@
 Validation through convergence diagnostics, posterior predictive checks, and residual analysis confirms the model's robustness and reliability within its assumptions.
-
-All analyses were conducted using R (@citeR) with the following key packages:
-
-- **tidyverse:** For data manipulation and visualization. (@tidyverse)
-- **rstanarm:** For Bayesian model fitting using Stan. (@rstanarm1, @rstanarm2)
-- **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
-- **bayesplot:** For comprehensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
 
@@
 ## Deep Red/Blue State Voting History {#sec-voting-hist}
+```{r}
+#| fig-cap: Number of Wins in 1000 Simulations by State
+#| echo: false
+#| warning: false
+#| message: false
 
-```{r}
 #| fig-cap: Voting History of the 10 States That do not Have Poll Data
@@
 #| message: false
+
+library(kableExtra)
+library(dplyr)
 
@@
 
-# -----------------------------
-# Step 3: Reshape the Data for a Compact Table
-# -----------------------------
-
 # Pair the states into two columns per row
-# We'll split the data into two halves
 df1 <- df[1:5, ]  # First 5 states
@@
 
-# -----------------------------
-# Step 4: Apply Cell-wise Coloring Based on Party
-# -----------------------------
-
-# Function to assign colors based on party
+# Function to assign color names based on party
 assign_color <- function(party) {
@@
   mutate(
-    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white"),
-    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white"),
-    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white"),
-    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white")
+    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
+    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
+    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE),
+    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE)
   )
-
-# -----------------------------
-# Step 5: Create the Styled Compact Table with Note
-# -----------------------------
 
@@
   select(State_1, Year_1, State_2, Year_2) %>%
-  kable("html", escape = FALSE, align = "c",
+  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c",
         col.names = c("State", "Last Time Voted for a Different Party",
                       "State", "Last Time Voted for a Different Party")) %>%
-  kable_styling("striped", full_width = FALSE, position = "left") %>%
-  # Header styling
-  row_spec(0, bold = TRUE, background = "#D3D3D3") %>%
-  # Add a footnote below the table
-  add_footnote("Note: Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.",
-  )
+  kable_styling(latex_options = c("striped"), full_width = FALSE, position = "left") %>%
+  row_spec(0, bold = TRUE) %>%
+  footnote(general = " Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.")
 
@@
 styled_table
+
 ```

### 2024-11-04 paper/paper.qmd
added data validation seciton for ideal poll

--- 2024-11-04 paper/paper.qmd
@@
 In an ideal world, where a $100k in funds and all necessary resources could be obtained from a neutral source for the purpose of polling the population, the idealized methodology for a survey that aims to forecast the results of the US election should be one that uses quota sampling and the probability proportionate to size method to target the states of interest specifically.
-Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past eight elections [@blueredstate].
-Out of the 51 states, there are 20 states which voted for the same party each time in the past nine elections and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
+Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past nine elections [@blueredstate].
+Out of the 51 states, there are 43 states which have voted for the same party seven or more times and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
+
 On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections—these would be the states of greater interest to us [@blueredstate].
@@
 This targetted sampling process allows us to focus our polling efforts on gathering more data from the areas where there is higher uncertainty, but the tradeoff is bias is easily introduced with the assumptions we are making, potentially impacting the accuracy of our forecast [@quotasampling].
-This means that more work is required to adjust for these biases while also minimizing the inflation of variance [@quotasamping].
+This means that more work is required to adjust for these biases while also minimizing the inflation of variance [@quotasampling].
 
@@
 
+When analyzing the survey responses, the Likert scale responses will be converted to a numerical scale for ease of analysis.
+During data validation, the collected responses should be checked to filter out any responses that may not reflect the participant's opinion accurately.
+
+For example, the logic between different questions should be verified to eliminate responses which clearly were not carefully thought through or misunderstood.
+For example, if the same respondent selected that they would vote for Kamala Harris and are very likely to continue supporting her, yet also indicate a lack of confidence in her abilities and lack of support for her within their close circle, then there is a logical conflict that may be caused by improper reading of the survey.
+As well, responses that are entirely "Don't know" selections for every single question should also be omitted from analysis, as these are likely participants who did not even attempt to read through the survey.
+
+\newpage
+
 ## Table Detailing First Few Observations of Dataset {#sec-app-table}
@@
 ```{r}
-#| label: tbl-1
+#| label: tbl-firstlines
 #| tbl-cap: First few observations of cleaned data set
@@
 
-```{=latex}
-\newpage\newpage
-
-```
 \newpage

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 
-- questions (wording, order, options)
-- treatment of non-response
-- offer of incentive
+The wording of questions should be straightforward and easy to understand, minimizing error in measurement caused by misinterpretation of questions.
+The question should also be asked neutrally, without leading participants to any particular response.
+Likert scales, including an option for “don’t know”, should be used whenever possible for ease and consistency of participant understanding.
+As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB].
+The “don’t know” option is also effective in helping reduce the decision difficulty to certain participants who are truly unsure.
+Before officially deploying the survey with our selected sample, it first needs to be tested to catch ambiguous or other poorly phrased questions [@citeTB].
+To encourage participation in this research, a financial incentive should be offered for full completion of the poll.
+Displaying the progress bar and indicating the survey length at the very beginning also helps encourage participants to complete the full questionnaire.
+Our poll would also be submitted to the appropriate ethics review boards to increase confidence in our methodology.
 
-After analysis of poll responses, the votes of the consistent states should be added back into consideration before making the final forecast for election results.
+Some information of interest to us, such as voting history and racial identification, is already included in the registered voter file, so these questions can be excluded from the poll itself [@voterfiledata].
+After analysis of poll responses, the votes of the historically consistent states should be added back into consideration before making the final forecast for election results.
 
 A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
-In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, such as Survey Monkey, the question order would be randomized when being presented to different respondents with the Question Randomization feature [@surveymonkey].
+In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, the question order would be randomized when being presented to different respondents, with consideration for related questions must be kept together.
 This will help reduce response order bias, which is when the answer of a question is affected by the order in which it appeared in the survey [@citeStantcheva].
+
+\newpage
 

### 2024-11-04 paper/paper.qmd

@@
 
+When analyzing the survey responses, the Likert scale responses will be converted to a numerical scale for ease of analysis.
+During data validation, the collected responses should be checked to filter out any responses that may not reflect the participant's opinion accurately.
+
+For example, the logic between different questions should be verified to eliminate responses which clearly were not carefully thought through or misunderstood.
+For example, if the same respondent selected that they would vote for Kamala Harris and are very likely to continue supporting her, yet also indicate a lack of confidence in her abilities and lack of support for her within their close circle, then there is a logical conflict that may be caused by improper reading of the survey.
+As well, responses that are entirely "Don't know" selections for every single question should also be omitted from analysis, as these are likely participants who did not even attempt to read through the survey.
+
 \newpage
@@
 
+=======
+>>>>>>> 3831a5556d47b44b38240855a1f7d590aa816a2d
 \newpage

### 2024-11-04 paper/paper.qmd
added critique of fox poll length

--- 2024-11-04 paper/paper.qmd
@@
 Additionally, enlisting high-profile figures or celebrities, like Elon Musk or LeBron James—who have recently endorsed Trump and Harris, respectively—could enhance their appeal in these swing states.
-These endorsements would help candidates reach diverse and niche demographics, broadening teheir voter base and potentially swaying undecided voters.
+These endorsements would help candidates reach diverse and niche demographics, broadening their voter base and potentially swaying undecided voters.
 
@@
 Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
-Some papers have included data from these markets in their models to improve their prediction accuracy.[REF https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+Some papers have included data from these markets in their models to improve their prediction accuracy [REF https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
 
@@
 This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” [@fox].
+Additionally, another potential weakness to this poll is its length.
+With around 50 questions, this survey will cause a large number of participants to lose interest before completion.
+This likely introduces high attrition of respondents dropping out mid-way, or measurement errors from respondents spending decreasing effort to think through their responses in the later presented questions [@surveylength].
 
@@
 Out of the 51 states, there are 43 states which have voted for the same party seven or more times and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
-
 On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections—these would be the states of greater interest to us [@blueredstate].

### 2024-11-04 paper/paper.qmd
added cites

--- 2024-11-04 paper/paper.qmd
@@
 Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
-Some papers have included data from these markets in their models to improve their prediction accuracy.[REF https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+Some papers have included data from these markets in their models to improve their prediction accuracy (@market_paper).
 
@@
 Furthermore, because these markets have only recently become popular or even legal in some cases one might expect these markets to be full of amateurs.
-These markets may display a bais as well, where the largest market Polymarket [REF] is based on crypto-currency, and Trump is popular among crypto-enthusiasts, and large financial institutions are underrepresented in the crypto-market.
+These markets may display a bais as well, where the largest market Polymarket is based on crypto-currency, and Trump is popular among crypto-enthusiasts, and large financial institutions are underrepresented in the crypto-market.
 
@@
 
-=======
->>>>>>> 3831a5556d47b44b38240855a1f7d590aa816a2d
 \newpage

### 2024-11-04 scripts/02-download_data.R

@@ -20,6 +20,8 @@ devtools::install_github("jaytimm/PresElectionResults")
 
 
 #### Download data ####
+#https://github.com/jaytimm/PresElectionResults
+
 historical_state_data <- PresElectionResults::pres_by_state
 
 

### 2024-11-04 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -11,9 +11,6 @@ library(tidyverse)
 
 #### Simulate data #### 
 
-#TODO:
-#Make simulations reflect close to reality, for example Harris will definitly win Hawaii, and Trump 
-#will win Kansas
 
 # set up vectors where we will sample from
 states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 

### 2024-11-04 paper/paper.qmd
resolved conflict

@@
 
-Every four years, the United States presidential election occurs and is one of the most significant political events internationally.
-The complex process of the election involves both the popular and electoral college system.
+Every four years, the United States presidential election is one of the most significant international political events.
+The complex election process involves both the popular and electoral college system.
 The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege].
@@
 
-Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs.
-For example, regarding technology, both parties are concerned with the rapid progression of artificial intelligence’s capabilities but differ in the response and regulation of it.
-Republicans believe that the current practices are unfair and disproportionately target their voices, whereas democrats generally are in favor of tighter regulation [@keyissues].
+Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs.
+For example, regarding technology, both parties are concerned with the accelerated progression of artificial intelligence’s capabilities but differ in the response and regulation of it.
+Republicans believe that the current practices are unfair and disproportionately target their voices, whereas Democrats generally are in favor of tighter regulation [@keyissues].
 Another key issue important to voters and the candidates is abortion.
-As former president, he supported restrictions on abortion access and advocated for policies that align with the conservative movement’s opposition to abortion rights [@abortionstance].
-On the other hand, Harris has been a key supporter in the pro-choice movement and has been an advocate for protecting and expanding access to abortion services, as well as female autonomy [@harrisviews].
+As a former president, he supported restrictions on abortion access and advocated for policies that align with the conservative movement’s opposition to abortion rights [@abortionstance].
+On the other hand, Harris has been a key supporter of the pro-choice movement and has been an advocate for protecting and expanding access to abortion services, as well as female autonomy [@harrisviews].
 The outcome of this election will have profound consequences for various legislative policies and the future direction of the United States.
 
-To forecast the 2024 United States presidential election, a generalized linear model and Bayesian model were built.
+To forecast the 2024 United States presidential election, a generalized linear model and a Bayesian model were built.
 These models will help us predict who will win the election based on the electoral college system.
-A key finding from our model is that we predict Kamala Harris to win the presidential election with 265 electoral college votes.
+An important finding from our model is that we predict Kamala Harris wins the presidential election with 279 electoral college votes.
 This would be a historic win, as she would be the first woman to become US president, a significant milestone in American politics.
-One drawback of our model stems from representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
+One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
+More is discussed in @sec-disc-imp.
 
 This paper is broken down into various sections, including Data, Modeling, Results, Discussion, Conclusion, and Appendices.
-This paper uses data made available by FiveThirtyEight [@538data] which compiles individual polling surveys based on state and methodology.
-@sec-data explores the data, highlighting key aspects that may be useful to future policymakers or campaign strategists, as well as details the variables present in the dataset.
+This paper uses data [@538data] which compiles individual polling surveys based on state and methodology.
+@sec-data explores the data, highlighting key aspects that may be useful to future policymakers or campaign strategists, detailing the variables present in the dataset.
 @sec-model presents the models that were built and used to forecast the election.
-@sec-results details the conclusions of our model and @sec-discussion explores possible implications and insights of our conclusions.
+@sec-results details the conclusions of our model and @sec-discussion explores possible implications and insights.
 @sec-appendices include an idealized methodology and survey that we could hypothetically run, with the task of forecasting the US presidential election.
@@
 The polling data extracted utilize the "pool of polls" method, which combines multiple polling results into a single estimate.
-By aggregating data from several polls, this method aims to create a more stable and reliable measure of public opinion by reducing the impact of outliers, sampling error, and individual poll biases.
+By aggregating data from several polls, this method aims to create a more stable and reliable measure of public opinion by reducing the impact of outliers, sampling errors, and individual poll biases.
 This dataset is crucial to help us build a model to forecast the 2024 US Presidential election and can help us analyze key features in voter habits for future political analysis.
 Note that polling data collection stopped on October 19.
-Our model’s results are based solely on data available as of that date and do not incorporate any subsequent polling updates.
+Our model’s results are based solely on data available as of that date and do not incorporate subsequent polling updates.
 
-There were two other potential datasets that could have been used, titled “Presidential Polling Averages” and “President Primary Polls”, both found through FiveThirtyEight [@538].
+Two other potential datasets could have been used, titled “Presidential Polling Averages” and “President Primary Polls”, both found through FiveThirtyEight [@538].
 The “Presidential Polling Averages” dataset only included data from the Biden versus Trump election in 2020 and was not updated to include data for the upcoming 2024 election and therefore inadequate.
-The “Presidential Primary Polls” dataset included very similar information as our chosen dataset, however, it included candidates that have since dropped out or were not major running candidates, such as Nikki Haley and Michelle Obama.
-It includes observations that are unnecessary and creates noise, therefore we did not choose this dataset either.
+The “Presidential Primary Polls” dataset included very similar information as our chosen dataset, however, it included candidates that have since dropped out or were not major running candidates, such as Nikki Haley and Michelle Obama, unnecessary observations, and therefore we did not choose this dataset either.
 
@@
 
-For further analysis, the cleaned dataset was separated into two different datasets by beach name, for simplicity purposes.
 Data was analyzed through the R programming software [@citeR] and packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], and dplyr [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
+For model construction and analysis, the following packages were used: rstanarm [@rstanarm1, @rstanarm2], brms [@brms1, @brms2, @brms3], and bayesplot [@bayesplot1, @bayesplot2]
 
@@
 
-Finally, @fig-3 shows the number of votes per candidate.
+Finally, @fig-3 shows the number of votes per candidate, based on popular vote.
 As expected, Harris and Trump have the highest number of votes, with Harris having slightly more votes based on the surveyed polls than Trump.
@@
 #| label: fig-3
-#| fig-cap: Number of Votes per Candidate. This reflects the popular vote for each candidate and does not account for the electoral college sytem. 
+#| fig-cap: Number of Votes per Candidate. This reflects the popular vote for each candidate and does not account for the electoral college system. 
 #| echo: false
@@
 - **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
-- **bayesplot:** For comprehensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
+- **bayesplot:** For extensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
 
@@
   *Definition:* Captures state-specific effects as fixed effects.  
-  *Justification:* Allows the model to account for regional variations in support that are not explained by other predictors, incorporating state-specific nuances.
+  *Justification:* Allows the model to account for regional variations in support that are not explained by other predictors, incorporating state-specific variations.
 
@@
 However, given the focus on specific state-level predictions and the availability of sufficient data per state, fixed effects are chosen in this case.
-Ideally we would fit both models and use diagnostics to choose which has lower bais and variance, and thus better predictive quality.
+Ideally, we would fit both models and use diagnostics to choose which has lower bias and variance for better predictive quality.
 Additionally, simpler models excluding variables like Transparency Score and Poll Score were evaluated but found inadequate in capturing the state-by-state differences in the transparency and reliability of the polls, potentially leading to biased support estimates.
@@
 
-Averaging predictors per state smooths out poll-to-poll variability and provides a representative set of predictors for each state, allowing use to make reliable state-level predictions.
+Averaging predictors per state smooths out poll-to-poll variability and provides a representative set of predictors for each state, allowing us to make reliable state-level predictions.
 Instead of averaging, a training/testing split or regional aggregations based on 538's political regions could have been employed.
@@
 
-Using the averaged state level predictors, we generated posterior predictions by drawing 1,000 samples from the posterior distributions of the model parameters.
+Using the averaged state-level predictors, we generated posterior predictions by drawing 1,000 samples from the posterior distributions of the model parameters.
 These predictions simulate 1,000 possible election outcomes, and by taking the proportion of victories for Harris and Trump we estimate the probability each candidate has of winning the election.
-To generate an election map, we simply take the average outcome of each state across the 1,000 simulations.
-Alternative methods could include using point estimates from the posterior mean; however, simulating multiple outcomes offers a more comprehensive assessment of uncertainty and variability in election outcomes.
+To generate an election map, we take the average outcome of each state across the 1,000 simulations.
+Alternative methods could include using point estimates from the posterior mean; however, simulating multiple outcomes offers a more extensive assessment of uncertainty and variability in election outcomes.
 
 For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
-This data was collected from a github repository (@hist_state) who collected their data from Wikipedia.
-This decision is based on the assumption that certain states are very unlikely to change their voting patterns, reducing the necessity for current polling data.
-The ten states that do not have polling data have not changed parties in at least the last 20 plus years (@sec-voting-hist).
+This data was collected from a GitHub repository (@hist_state) which collected their data from Wikipedia.
+This decision is based on the assumption that certain states are unlikely to change their voting patterns, reducing the necessity for current polling data.
+The ten states that do not have polling data have not changed parties in at least the last 20-plus years (@sec-voting-hist).
 Alternatively, imputation methods or political regional averages could have been used, but historical averages are used for their simplicity and relevance.
 
-Though Maine and Nebraska allow for their Electoral College votes to be split.
+Maine and Nebraska allow for their Electoral College votes to be split.
 In Maine two of its four votes are awarded statewide and the other two for each of its congressional districts.
 In Nebraska two of its five votes are awarded statewide and the other three for each of its congressional districts.
-Our model simplifies this dynamic and all of their votes are assigned according the support overall in the state.
+Our model simplifies this dynamic and all of their votes are assigned according to the support overall in the state.
 The main reason for this is that some of the districts have little or no polling data, making it difficult to measure their support for either candidate.
@@
 
-Comprehensive diagnostics were performed to assess model convergence, fit, and predictive performance.
+Extensive diagnostics were performed to assess model convergence, fit, and predictive performance.
 Graphs and more details can be found in the relevant appendices.
 We evaluated model convergence using the R-hat statistic and trace plots.
-Trace plots, @sec-trace-plots revealed well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process.
+Trace plots, @sec-trace-plots show well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process.
 
@@
 
-Residual plots (found in @sec-residual-plots, depicting residuals versus fitted values, showed notable clumping in the central region for both models, along with vertical lines.
+Residual plots (found in @sec-residual-plots, depicting residuals versus fitted values) showed notable clumping in the central region for both models, along with vertical lines.
 This pattern may indicate unmodeled heterogeneity or data limitations, suggesting areas where the model's fit could be improved.
-Despite these observations, the overall residual distribution did not reveal significant systematic errors, supporting the model's general adequacy.
+Despite these observations, the overall residual distribution did not show significant systematic errors, supporting the model's general adequacy.
 
-Ideally several models would be created then using model selection techniques would be used to determine the best model for predicting the outcome of the US election.
-Due to time constraints the following models were considered but not implemented.
+Ideally, multiple models would be created.
+Then, model selection techniques would be used to determine the best model for predicting the outcome of the US election.
+Due to time constraints, the following models were considered but not implemented.
 A set of Bayesian models with state as a random effect and models excluding specific predictors like Transparency Score.
@@
 
-@table-2 displays each candidates chance of winning and average number of Elector Votes they received over the 1000 simulated elections.
+@table-2 displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
 Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
@@
 Based on the simulated elections, @fig-5 illustrates the most likely winner of each state.
-In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.
+In this scenario, Harris wins 276 Electoral Votes, and Trump wins 262.
 A table in @sec-state-wins, shows how this map was constructed by taking the most probable winners of each state.
@@
 The results of the US elections in large part hinge on the direction of the swing states.
-A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris.
+A swing state is a state that holds a significant number of electoral votes and could reasonably be won by either Trump or Harris.
 @fig-6 shows the probability that both candidates have of winning each of the 7 swing states.
@@
 One of the primary challenges stems from gaps in our dataset, specifically the absence of polling data for several states.
-When the polling data was collected and compiled by @538, it did not comprehensively include every state, leading to missing data points that affect our predictions.
+When the polling data was collected and compiled by @538, it did not extensively include every state, leading to missing data points that affected our predictions.
 Our model relies on the electoral college system, where each state’s popular vote determines its allocation of electoral votes, ultimately deciding the election’s outcome.
@@
 
-Our model does not take into account state correlations based on past election data and research based insights into voter preference.
-For example a model developed by Chernov, Elenev, and Song, posits if Harris were to win Pennsylvania, then her chances of winning Nevada would increase from 24% to 68% [https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
-These correlations between states could have a large impact on election predictions and our model is limited by not modelling these interactions.
+Our model does not take into account state correlations based on past election data and research-based insights into voter preference.
+For example, a model developed by Chernov, Elenev, and Song, posits if Harris were to win Pennsylvania, then her chances of winning Nevada would increase from 24% to 68% [https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+These correlations between states could have a large impact on election predictions and our model is limited by not modeling these interactions.
 
@@
 Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
-Some papers have included data from these markets in their models to improve their prediction accuracy [REF https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+Some papers have included data from these markets in their models to improve their prediction accuracy (@market_paper).
 
-Some key advantages of these markets is that they are real time, so the market adjusts quickly to important and material news.
-Where as poll are sluggish to incorporate big changes as new polls need to be sent out and analysed.
-Another advantage is that these markets may attract sophisticated investors that could employ methods that differ from polling but still offer predictive capability.
+Some key advantages of these markets is that they are real-time, so the market adjusts quickly to important and material news, whereas polls are sluggish to incorporate big changes, since new polls need to be sent out and analyzed.
+Another advantage is that these markets may attract sophisticated investors who could employ methods that differ from polling but still offer predictive capability.
 The larger the trade volume the more likely these actors are to be present and wash out any unsophisticated trading.
 
-On the other hand the main disadvantage is that all of the investors in this market essentially completely anonymous.
-This means there is no way of checking whether a large investor, who has significant price setting power, is indeed sophisticated or not.
-Furthermore, because these markets have only recently become popular or even legal in some cases one might expect these markets to be full of amateurs.
-These markets may display a bais as well, where the largest market Polymarket [REF] is based on crypto-currency, and Trump is popular among crypto-enthusiasts, and large financial institutions are underrepresented in the crypto-market.
+On the other hand, the main disadvantage is that all investors in this market are completely anonymous.
+This means there is no way of checking whether a large investor, who has significant price-setting power, is indeed sophisticated or not.
+Furthermore, because these markets have only recently become popular, or even legal in some cases, one might expect these markets to be full of amateurs.
+These markets may display bias as well.
+For example, the largest market, Polymarket, is based on crypto-currency.
+The bias lies in the fact that Trump is popular among crypto-enthusiasts and large financial institutions are underrepresented in the crypto-market.
 
-This is a new frontier in forecasting and including data from these markets may prove beneficial. Or these markets may be nothing more than a hobby, and hold no useful data.
+This is a new frontier in forecasting and including data from these markets may prove beneficial. Or these markets may be nothing more than a hobby and hold no useful data.
 
@@
 One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research [@fox].
-Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phonecall interviews or an online survey.
+Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phone interviews or an online survey.
 The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers [@fox].
 This sampling method is a type of probability sampling in which each unit's chance of being included is dependent upon their size, a measure that is based off of some characteristic that is known about every single unit and often related to the main variable of interest.
-By giving a higher probability of inclusion to units of greater importance based on its size during sampling, a more accurate estimate can be obtained [@Latpate2021].
+By giving a higher probability of inclusion to units of greater importance based on their size during sampling, a more accurate estimate can be obtained [@Latpate2021].
 In the case of this poll, the number of voters per state region was used to determine the proportion of individuals contacted [@fox].
@@
 One strength of this questionnaire is that information regarding the respondent’s confidence level and commitment level was also collected.
-With questions such as “How often do you make a point to read or listen to the news?” and “Are you certain to support that candidate, or do you think you may change your mind and support someone else?”, this poll is able to gain a more indepth view of how easily these participants may be swayed in the time between the polling and the real election  [@fox].
-As well, probabilistic statistical models based on past voting history, interest in current election, age, education, race, ethnicity, church attendance, and marital status were used to predict the respondents’ likeliness to vote; interestingly, results found that the results for the full sample vs subsample of 870 likely voters varied by a ±3%, which can be quite significant given how tight the race is currently  [@fox].
+With questions such as “How often do you make a point to read or listen to the news?” and “Are you certain to support that candidate, or do you think you may change your mind and support someone else?”, this poll is able to gain a more in-depth view of how easily these participants may be swayed in the time between the polling and the real election  [@fox].
+As well, probabilistic statistical models based on past voting history, interest in the current election, age, education, race, ethnicity, church attendance, and marital status were used to predict the respondents’ likeliness to vote; interestingly, results found that the results for the full sample vs subsample of 870 likely voters varied by a ±3%, which can be quite significant given how tight the race is currently  [@fox].
 
@@
 This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” [@fox].
-Additionally, another potential weakness to this poll is its length.
+Additionally, another potential weakness of this poll is its length.
 With around 50 questions, this survey will cause a large number of participants to lose interest before completion.
@@
 
-In an ideal world, where a $100k in funds and all necessary resources could be obtained from a neutral source for the purpose of polling the population, the idealized methodology for a survey that aims to forecast the results of the US election should be one that uses quota sampling and the probability proportionate to size method to target the states of interest specifically.
+In an ideal world, where $100k in funds and all necessary resources could be obtained from a neutral source to poll the population, the idealized methodology for a survey that aims to forecast the results of the US election should use quota sampling and the probability proportionate to size method to target the states of interest specifically.
 Historically, certain states have very strong and consistent party preferences; for example, we can be almost entirely certain even without conducting any current research that California would be voting Democrat, simply because this has been how they voted in the past nine elections [@blueredstate].
 Out of the 51 states, there are 43 states which have voted for the same party seven or more times and therefore there is a high probability that the pattern will be upheld with this year's election [@blueredstate].
-On the other end of the spectrum, there are seven states which have voted for each party three of more times within the last nine elections—these would be the states of greater interest to us [@blueredstate].
+On the other end of the spectrum, seven states have voted for each party three or more times within the last nine elections—these would be the states of greater interest to us [@blueredstate].
 
@@
 
-After the sample has been selected with the above methods, the poll itself will be conducted through a survey sent to the chosen participants in order to reduce interviewer bias, or the effect of the interviewer’s own views on the measured responses, which can affect results when face-to-face or phone interviews are utilized [@citeTB].
-As well, to reduce social desirability effect, the survey must emphasize in the beginning that the respondents’ identity will be kept anonymous to the researchers and the final reported data [@citeStantcheva].
+After the sample has been selected with the above methods, the poll itself will be conducted through a survey sent to the chosen participants to reduce interviewer bias, or the effect of the interviewer’s views on the measured responses, which can affect results when face-to-face or phone interviews are utilized [@citeTB].
+As well, to reduce the social desirability effect, the survey must emphasize in the beginning that the respondents’ identity will be kept anonymous to the researchers and the final reported data [@citeStantcheva].
 
-The wording of questions should be straightforward and easy to understand, minimizing error in measurement caused by misinterpretation of questions.
+The wording of questions should be straightforward to understand, minimizing measurement error, caused by misinterpretation of questions.
 The question should also be asked neutrally, without leading participants to any particular response.
@@
 As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB].
-The “don’t know” option is also effective in helping reduce the decision difficulty to certain participants who are truly unsure.
+The “don’t know” option is also effective in helping reduce the decision difficulty of certain participants who are truly unsure.
 Before officially deploying the survey with our selected sample, it first needs to be tested to catch ambiguous or other poorly phrased questions [@citeTB].
@@
 
-A sample implementation of such poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
-In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with more professional surveying platform, the question order would be randomized when being presented to different respondents, with consideration for related questions must be kept together.
-This will help reduce response order bias, which is when the answer of a question is affected by the order in which it appeared in the survey [@citeStantcheva].
+A sample implementation of such a poll, created using Google form, can be found [here](https://forms.gle/ti2rjGGVxBukv9dF6).
+In this version, the order of questions presented is consistent for all respondents due to the limitations of Google form, but for an idealized methodology with a more professional surveying platform, the question order would be randomized when being presented to different respondents, with consideration for related questions must be kept together.
+This will help reduce response order bias, which is when the answer to a question is affected by the order in which it appeared in the survey [@citeStantcheva].
 
@@
 
-For example, the logic between different questions should be verified to eliminate responses which clearly were not carefully thought through or misunderstood.
+For example, the logic between different questions should be verified to eliminate responses that were not carefully thought through or misunderstood.
 For example, if the same respondent selected that they would vote for Kamala Harris and are very likely to continue supporting her, yet also indicate a lack of confidence in her abilities and lack of support for her within their close circle, then there is a logical conflict that may be caused by improper reading of the survey.
-As well, responses that are entirely "Don't know" selections for every single question should also be omitted from analysis, as these are likely participants who did not even attempt to read through the survey.
+As well, responses that are entirely "Don't know" selections for every single question should also be omitted from the analysis, as these are likely participants who did not even attempt to read through the survey.
 

### 2024-11-04 paper/paper.qmd
added section on clinton effect

@@
 
+A further limitation of our model’s prediction is that there could be inherent biases present in the data which are difficult to locate due to the vast size of the compiled dataset.
+In this election, there is a greater demographic contrast between candidates than ever before, with a Black woman running against a White man.
+With such contrasting identities being represented, the respondents are more likely than ever to be affected by biases and prejudices based on the candidates’ race and gender.
+However, due to social desirability bias, many respondents may choose to not answer or answer more neutrally in order to conceal the strength of their disapproval for Harris due to these factors.
+This may cause a skew where predictions indicate significantly higher support for Harris than is actually present in the population.
+
+Despite survey predictions successfully minimizing the effect of social desirability when predicting the election results in 2008 and 2012 for Barack Obama, it had a significant effect in 2016 when the only other woman, Hillary Clinton, ran for president [@clintoneffect].
+The election that year saw that previously undecided voters largely leaned against Clinton, and finally ended with her loss [@clintoneffect].
+With this in mind, it is uncertain how much of the neutral and non-responses within the current dataset are caused by this bias, meaning that the higher support this current model predicts for Harris also could be a misleading result.
+
 ## Next Steps {#sec-disc-next}

### 2024-11-04 paper/paper.qmd
separated weaknesses into its own section

@@
 One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
-More is discussed in @sec-disc-imp.
+More is discussed in @sec-disc-weaknesses.
 
@@
 
-## Implications and Potential Weaknesses {#sec-disc-imp}
+## Result Implications {#sec-disc-imp}
 
@@
 However, the forecast indicates that Harris is expected to win by narrow margins of approximately 1.98%, 1.47%, and 1.46% in these states, respectively.
+
+## Potential Weaknesses {#sec-disc-weaknesses}
 
@@
 Our model does not take into account state correlations based on past election data and research-based insights into voter preference.
-For example, a model developed by Chernov, Elenev, and Song, posits if Harris were to win Pennsylvania, then her chances of winning Nevada would increase from 24% to 68% [https://anderson-review.ucla.edu/wp-content/uploads/2024/10/voter_preferences_updated.pdf].
+For example, a model developed by Chernov, Elenev, and Song, posits if Harris were to win Pennsylvania, then her chances of winning Nevada would increase from 24% to 68% [@market_paper].
 These correlations between states could have a large impact on election predictions and our model is limited by not modeling these interactions.

### 2024-11-04 scripts/06-bayes_model.R
added llm usage, removed plots

@@ -334,31 +334,6 @@ saveRDS(model_trump, file = "models/bayes_model_trump.rds")
 
 ### Model Diagnostics ###
 
-# Plot trace plots for Trump
-trump_trace <- plot(model_trump, plotfun = "trace")
-
-# Plot trace plots for Harris
-harris_trace <- plot(model_harris, plotfun = "trace")
-
-ggsave(
-  filename = "trump_trace.pdf",        # File name
-  plot = trump_trace,                  # Plot object
-  path = "other/plots",               # Directory to save (create if it doesn't exist)
-  width = 10,                            # Width in inches
-  height = 6,                            # Height in inches
-  dpi = 300                              # Resolution
-)
-
-
-ggsave(
-  filename = "harris_trace.pdf",        # File name
-  plot = harris_trace,                  # Plot object
-  path = "other/plots",               # Directory to save (create if it doesn't exist)
-  width = 10,                            # Width in inches
-  height = 6,                            # Height in inches
-  dpi = 300                              # Resolution
-)
-
 
 # PPC Checks
 

### 2024-11-04 paper/paper.qmd

@@
 
-Extensive diagnostics were performed to assess model convergence, fit, and predictive performance.
+Diagnostics were performed to assess model convergence, fit, and predictive performance.
 Graphs and more details can be found in the relevant appendices.
-We evaluated model convergence using the R-hat statistic and trace plots.
-Trace plots, @sec-trace-plots show well-mixed chains without discernible trends, further supporting the reliability of the MCMC sampling process.
-
 Posterior predictive checks were conducted to assess the model's ability to replicate observed data, @sec-ppc.
@@
 
-## Trace Plots {#sec-trace-plots}
-
-```{r}
-#| fig-cap: Trump Trace Plots
-#| echo: false
-#| warning: false
-#| message: false
-
-plot_path <- here::here("other", "plots","trump_trace.pdf" )
-knitr::include_graphics(plot_path)
-
-```
-
-```{r}
-#| fig-cap: Harris Trace Plots
-#| echo: false
-#| warning: false
-#| message: false
-
-plot_path <- here::here("other", "plots","harris_trace.pdf" )
-knitr::include_graphics(plot_path)
-
-```
-
 ## Posterior Predictive Checks {#sec-ppc}

### 2024-11-04 paper/paper.qmd
intext citatiaon formating
fixed up

@@
 ```
-The results of the US elections in large part hinge on the direction of the swing states.
-A swing state is a state that holds a significant number of electoral votes and could reasonably be won by either Trump or Harris.
+The outcome of the US election may be entirely determined by seven influential states, if the other 44 voting region vote as they are expected to.
+These are the swing states, these states hold a significant number of electoral votes and could reasonably be won by either Trump or Harris.
 @fig-6 shows the probability that both candidates have of winning each of the 7 swing states.
@@
 ```{r}
-#| fig-cap: Number of Wins in 1000 Simulations by State
-#| echo: false
-#| warning: false
-#| message: false
-
 #| fig-cap: Voting History of the 10 States That do not Have Poll Data

### 2024-11-04 scripts/05-model_data.R

@@ -1,227 +1,0 @@
-#### Preamble ####
-# Purpose: Models the regional support, then calculates the EV each canditate is
-#          expected to get
-# Author:  Robert Ford, Michelle Ji, Cher Ning
-# Date: 14 October 2024
-# Contact: robert.ford@mail.utoronto.ca, michelle.ji@mail.utoronto.ca, cher.ning@mail.utoronto.ca
-# License: MIT
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(tidyr)
-
-#### Read data ####
-region_mapping <- c(
-  # Pacific Region
-  "Washington" = "Pacific",
-  "Oregon" = "Pacific",
-  "California" = "Pacific",
-  "Hawaii" = "Pacific",
-  
-  # Mountain Region
-  "Alaska" = "Mountain",
-  "Idaho" = "Mountain",
-  "Montana" = "Mountain",
-  "Wyoming" = "Mountain",
-  "Utah" = "Mountain",
-  
-  # Southwest
-  "Nevada" = "Southwest",
-  "Colorado" = "Southwest",
-  "Arizona" = "Southwest",
-  "New Mexico" = "Southwest",
-  
-  # Plains
-  "North Dakota" = "Plains",
-  "South Dakota" = "Plains",
-  "Nebraska" = "Plains",
-  "Kansas" = "Plains",
-  
-  # Rust Belt
-  "Minnesota" = "Rust",
-  "Iowa" = "Rust",
-  "Wisconsin" = "Rust",
-  "Michigan" = "Rust",
-  "Illinois" = "Rust",
-  "Indiana" = "Rust",
-  "Ohio" = "Rust",
-  "Pennsylvania" = "Rust",
-  
-  # Texish
-  "Oklahoma" = "Texish",
-  "Texas" = "Texish",
-  "Louisiana" = "Texish",
-  
-  # South
-  "Missouri" = "South",
-  "Arkansas" = "South",
-  "Kentucky" = "South",
-  "Tennessee" = "South",
-  "Mississippi" = "South",
-  "Alabama" = "South",
-  "West Virginia" = "South",
-  "South Carolina" = "South",
-  
-  # Southeast
-  "Virginia" = "Southeast",
-  "North Carolina" = "Southeast",
-  "Georgia" = "Southeast",
-  "Florida" = "Southeast",
-  
-  # Northeast
-  "New York" = "Northeast",
-  "New Jersey" = "Northeast",
[…]

### 2024-11-04 paper/paper.qmd
syntax edits

--- 2024-11-04 paper/paper.qmd
@@
 
-Data was analyzed through the R programming software [@citeR] and packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], and dplyr [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
-For model construction and analysis, the following packages were used: rstanarm [@rstanarm1, @rstanarm2], brms [@brms1, @brms2, @brms3], and bayesplot [@bayesplot1, @bayesplot2]
+Data was analyzed through the R programming software [@citeR] and packages such as `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `knitr` [@knitr], and `dplyr` [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
+For model construction and analysis, the following packages were used: `rstanarm` [@rstanarm1, @rstanarm2], `brms` [@brms1, @brms2, @brms3], and `bayesplot` [@bayesplot1, @bayesplot2].
 
@@
 
-All model construction and analysis was conducted using R (@citeR) with the following key packages:
+All model construction and analysis was conducted using R [@citeR] with the following key packages:
 
-- **tidyverse:** For data manipulation and visualization. (@tidyverse)
-- **rstanarm:** For Bayesian model fitting using Stan. (@rstanarm1, @rstanarm2)
-- **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
-- **bayesplot:** For extensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
+- **tidyverse:** For data manipulation and visualization [@tidyverse].
+- **rstanarm:** For Bayesian model fitting using Stan [@rstanarm1; @rstanarm2].
+- **brms:** Alternative package for Bayesian regression modeling [@brms1; @brms2; @brms3].
+- **bayesplot:** For extensive model diagnostics and visualization [@bayesplot1; @bayesplot2].
 
@@
 For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
-This data was collected from a GitHub repository (@hist_state) which collected their data from Wikipedia.
+This data was collected from a GitHub repository [@hist_state] which collected their data from Wikipedia.
 This decision is based on the assumption that certain states are unlikely to change their voting patterns, reducing the necessity for current polling data.
@@
 A set of Bayesian models with state as a random effect and models excluding specific predictors like Transparency Score.
-More complex models with additional interaction terms or non-linear components were also considered
+More complex models with additional interaction terms or non-linear components were also considered.
 
@@
 
-@table-2 displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
+@fig-predic displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
 Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
@@
 ```{r}
-#| label: table-2
+#| label: fig-predic
 #| fig-cap: Election Prediction Results
@@
 Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
-Some papers have included data from these markets in their models to improve their prediction accuracy (@market_paper).
+Some papers have included data from these markets in their models to improve their prediction accuracy [@market_paper].
 

### 2024-11-04 paper/paper.qmd
citation

@@
 Data was analyzed through the R programming software [@citeR] and packages such as `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `knitr` [@knitr], and `dplyr` [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
-For model construction and analysis, the following packages were used: `rstanarm` [@rstanarm1, @rstanarm2], `brms` [@brms1, @brms2, @brms3], and `bayesplot` [@bayesplot1, @bayesplot2].
+For model construction and analysis, the following packages were used: `rstanarm` [@rstanarm1; @rstanarm2], `brms` [@brms1; @brms2; @brms3], and `bayesplot` [@bayesplot1; @bayesplot2].
 

### 2024-11-04 paper/paper.qmd
added fig 6 caption

@@
 #| label: fig-6
+#| fig-cap: Probability of Each Candidate Winning in the Swing States
 #| echo: false

### 2024-11-04 paper/paper.qmd
fixed typos

--- 2024-11-04 paper/paper.qmd
@@
 ---
-title: "Forecasting the United States 2024 Presidential Election"
+title: "Forecasting the United States 2024 Presidential Election Using State as Fixed Effect Through a Bayesain Approach"
 subtitle: "Kamala Harris wins 60th presidency with 279 Electoral College Votes"
@@
 The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege].
-After the popular vote is accounted for, the candidate who receives the majority of the electoral votes from all states wins the presidential election.
-The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege].
+After the popular vote is accounted for, the candidate who receives the majority of the electoral votes wins the presidential election.
+The number of electoral votes is proportional to each state’s respective population [@electoralcollege].
 
@@
 For example, regarding technology, both parties are concerned with the accelerated progression of artificial intelligence’s capabilities but differ in the response and regulation of it.
-Republicans believe that the current practices are unfair and disproportionately target their voices, whereas Democrats generally are in favor of tighter regulation [@keyissues].
+Republicans believe that current moderation practices are unfair and disproportionately target their voices, whereas Democrats generally are in favor of tighter regulation [@keyissues].
 Another key issue important to voters and the candidates is abortion.
@@
 
-To forecast the 2024 United States presidential election, a generalized linear model and a Bayesian model were built.
+To forecast the 2024 United States presidential election a Bayesian model was built.
 These models will help us predict who will win the election based on the electoral college system.
@@
 This would be a historic win, as she would be the first woman to become US president, a significant milestone in American politics.
-One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
+One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls (after the cleaning process), which may impact the our estimates of electoral college votes.
 More is discussed in @sec-disc-weaknesses.
@@
 
-The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
+The variables used in the dataset include pollster name, pollster weight, method, state, sample size, Poll Score, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
 Pollster Name indicates which pollster the individual poll came from and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time to adjust for any bias.
@@
 State indicates what state the poll was conducted and sample size is the size of the poll.
-The pollscore is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number.
+The Poll Score is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number.
 The election date and days to election variables indicate how far out the poll was conducted with respect with the election date of November 5, 2024.
@@
 
-Data was analyzed through the R programming software [@citeR] and packages such as `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `knitr` [@knitr], and `dplyr` [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
-For model construction and analysis, the following packages were used: `rstanarm` [@rstanarm1; @rstanarm2], `brms` [@brms1; @brms2; @brms3], and `bayesplot` [@bayesplot1; @bayesplot2].
+Data was analyzed through the R programming software [@citeR] and packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], and dplyr [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
+For model construction and analysis, the following packages were used: rstanarm [@rstanarm1, @rstanarm2], brms [@brms1, @brms2, @brms3], and bayesplot [@bayesplot1, @bayesplot2]
 
@@
 
-All model construction and analysis was conducted using R [@citeR] with the following key packages:
+All model construction and analysis was conducted using R (@citeR) with the following key packages:
 
-- **tidyverse:** For data manipulation and visualization [@tidyverse].
-- **rstanarm:** For Bayesian model fitting using Stan [@rstanarm1; @rstanarm2].
-- **brms:** Alternative package for Bayesian regression modeling [@brms1; @brms2; @brms3].
-- **bayesplot:** For extensive model diagnostics and visualization [@bayesplot1; @bayesplot2].
+- **tidyverse:** For data manipulation and visualization. (@tidyverse)
+- **rstanarm:** For Bayesian model fitting using Stan. (@rstanarm1, @rstanarm2)
+- **brms:** Alternative package for Bayesian regression modeling. (@brms1, @brms2, @brms3)
+- **bayesplot:** For extensive model diagnostics and visualization.(@bayesplot1, @bayesplot2)
 
@@
 For states lacking recent polling data, we incorporated historical averages of Democratic and Republican support from elections since 2000.
-This data was collected from a GitHub repository [@hist_state] which collected their data from Wikipedia.
+This data was collected from a GitHub repository (@hist_state) which collected their data from Wikipedia.
 This decision is based on the assumption that certain states are unlikely to change their voting patterns, reducing the necessity for current polling data.
@@
 A set of Bayesian models with state as a random effect and models excluding specific predictors like Transparency Score.
-More complex models with additional interaction terms or non-linear components were also considered.
+More complex models with additional interaction terms or non-linear components were also considered
 
-#### Assumptions and Limitations
+**Assumptions and Limitations**
 
@@
 
-@fig-predic displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
+This table displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
 Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
@@
 ```{r}
-#| label: fig-predic
+#| label: table-2
 #| fig-cap: Election Prediction Results
@@
 Recently the trading volume on these markets is significant, which indicates that these markets may be helpful in predicting election outcomes.
-Some papers have included data from these markets in their models to improve their prediction accuracy [@market_paper].
+Some papers have included data from these markets in their models to improve their prediction accuracy (@market_paper).
 
@@
 
-library(kableExtra)
-library(dplyr)
-
-# Create a data frame with the provided data
+# Create the data frame
 df <- data.frame(
@@
 
-# Combine into a single data frame with 6 columns
+# Combine into a single data frame with 4 columns
 compact_df <- data.frame(
@@
   Year_1 = df1$Start_Year,
-  Party_1 = df1$Party,
   State_2 = df2$State,
   Year_2 = df2$Start_Year,
-  Party_2 = df2$Party,
   stringsAsFactors = FALSE
@@
 
-# Function to assign color names based on party
-assign_color <- function(party) {
-  ifelse(party == "Rep", "#DD1717", "#0F4392")
-}
+# Load the knitr package
+library(knitr)
 
-# Apply cell_spec to each cell that needs coloring
-compact_df <- compact_df %>%
-  mutate(
-    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
-    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
-    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE),
-    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE)
-  )
-
-# Create and style the table
-styled_table <- compact_df %>%
-  select(State_1, Year_1, State_2, Year_2) %>%
-  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c",
-        col.names = c("State", "Last Time Voted for a Different Party",
-                      "State", "Last Time Voted for a Different Party")) %>%
-  kable_styling(latex_options = c("striped"), full_width = FALSE, position = "left") %>%
-  row_spec(0, bold = TRUE) %>%
-  footnote(general = " Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.")
-
-# Display the table
-styled_table
-
+# Create the table
+kable(compact_df, col.names = c("State", "Last Time Voted for a Different Party",
+                                "State", "Last Time Voted for a Different Party"))
 ```

### 2024-11-04 paper/paper.qmd

@@
 
-Data was analyzed through the R programming software [@citeR] and packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], and dplyr [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
-For model construction and analysis, the following packages were used: rstanarm [@rstanarm1, @rstanarm2], brms [@brms1, @brms2, @brms3], and bayesplot [@bayesplot1, @bayesplot2]
+Data was analyzed through the R programming software [@citeR] and packages such as `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `knitr` [@knitr], and `dplyr` [@dplyr] were used to help download, clean, simulate, analyze, and test the data.
+For model construction and analysis, the following packages were used: `rstanarm` [@rstanarm1; @rstanarm2], `brms` [@brms1; @brms2; @brms3], and `bayesplot` [@bayesplot1; @bayesplot2].
 
@@
 There is also a 0.9% chance of a tie On average across the 1,000 simulations Harris wins 279.33 Electoral Votes, while Trump wins on average 258.67 votes.
+@fig-predic displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
+Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
+While Trump only has a 31.7% chance of winning.
+There is also a 0.9% chance of a tie On average across the 1,000 simulations Harris wins 279.33 Electoral Votes, while Trump wins on average 258.67 votes.
 
 ```{r}
-#| label: table-2
+#| label: fig-predic
 #| fig-cap: Election Prediction Results
@@
 
-# Load the knitr package
-library(knitr)
-
 # Create the table

### 2024-11-04 paper/paper.qmd
typo

@@
 - **Support**  
-  *Definition:* The support level for candidate \( i \) in a given poll.  
-  *Justification:* Measures the proportion of respondents supporting candidate \( i \), providing insight into their current standing.
+  *Definition:* The support level for candidate in a given poll.  
+  *Justification:* Measures the proportion of respondents supporting candidate, providing insight into their current standing.
 
 - **Sample Size**  
-  *Definition:* Represents the number of respondents in poll \( i \).  
+  *Definition:* Represents the number of respondents in poll.  
   *Justification:* Larger sample sizes generally yield more accurate estimates of support, reducing sampling variability.
@@
 - **Days to election**  
-  *Definition:* Denotes the number of days remaining until the election when poll \( i \) was conducted.  
+  *Definition:* Denotes the number of days remaining until the election when poll was conducted.  
   *Justification:* Accounts for temporal proximity to the election, capturing potential fluctuations in support as the election approaches.
@@
 
-where \(N_s\) is the number of polls for state \(s\).
+where \(N\) is the number of polls for state.
 

### 2024-11-04 paper/paper.qmd
removed unnecesary new page

--- 2024-11-04 paper/paper.qmd
@@
 ---
-title: "Forecasting the United States 2024 Presidential Election Using State as Fixed Effect Through a Bayesain Approach"
+title: "Forecasting the United States 2024 Presidential Election"
 subtitle: "Kamala Harris wins 60th presidency with 279 Electoral College Votes"
@@
 The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege].
-After the popular vote is accounted for, the candidate who receives the majority of the electoral votes wins the presidential election.
-The number of electoral votes is proportional to each state’s respective population [@electoralcollege].
+After the popular vote is accounted for, the candidate who receives the majority of the electoral votes from all states wins the presidential election.
+The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege].
 
@@
 For example, regarding technology, both parties are concerned with the accelerated progression of artificial intelligence’s capabilities but differ in the response and regulation of it.
-Republicans believe that current moderation practices are unfair and disproportionately target their voices, whereas Democrats generally are in favor of tighter regulation [@keyissues].
+Republicans believe that the current practices are unfair and disproportionately target their voices, whereas Democrats generally are in favor of tighter regulation [@keyissues].
 Another key issue important to voters and the candidates is abortion.
@@
 
-To forecast the 2024 United States presidential election a Bayesian model was built.
+To forecast the 2024 United States presidential election, a generalized linear model and a Bayesian model were built.
 These models will help us predict who will win the election based on the electoral college system.
@@
 This would be a historic win, as she would be the first woman to become US president, a significant milestone in American politics.
-One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls (after the cleaning process), which may impact the our estimates of electoral college votes.
+One drawback of our model stems from the representativeness of the dataset, as it does not include all 50 states in the polls, which may impact the electoral college votes.
 More is discussed in @sec-disc-weaknesses.
@@
 
-The variables used in the dataset include pollster name, pollster weight, method, state, sample size, Poll Score, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
+The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region.
 Pollster Name indicates which pollster the individual poll came from and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time to adjust for any bias.
@@
 State indicates what state the poll was conducted and sample size is the size of the poll.
-The Poll Score is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number.
+The pollscore is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number.
 The election date and days to election variables indicate how far out the poll was conducted with respect with the election date of November 5, 2024.
@@
 
-**Assumptions and Limitations**
+#### Assumptions and Limitations
 
@@
 
-This table displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
-Our model predicts that Harris has a 67.4% chance of winning the US presidential election.
-While Trump only has a 31.7% chance of winning.
-There is also a 0.9% chance of a tie On average across the 1,000 simulations Harris wins 279.33 Electoral Votes, while Trump wins on average 258.67 votes.
 @fig-predic displays each candidate's chance of winning and the average number of Elector Votes they received over the 1000 simulated elections.
@@
 ```
-The outcome of the US election may be entirely determined by seven influential states, if the other 44 voting region vote as they are expected to.
-These are the swing states, these states hold a significant number of electoral votes and could reasonably be won by either Trump or Harris.
+The results of the US elections in large part hinge on the direction of the swing states.
+A swing state is a state that holds a significant number of electoral votes and could reasonably be won by either Trump or Harris.
 @fig-6 shows the probability that both candidates have of winning each of the 7 swing states.
@@
 
-\newpage
-
 ## Table Detailing First Few Observations of Dataset {#sec-app-table}
@@
 ```{r}
+#| fig-cap: Number of Wins in 1000 Simulations by State
+#| echo: false
+#| warning: false
+#| message: false
+
 #| fig-cap: Voting History of the 10 States That do not Have Poll Data
@@
 
-# Create the data frame
+library(kableExtra)
+library(dplyr)
+
+# Create a data frame with the provided data
 df <- data.frame(
@@
 
-# Combine into a single data frame with 4 columns
+# Combine into a single data frame with 6 columns
 compact_df <- data.frame(
@@
   Year_1 = df1$Start_Year,
+  Party_1 = df1$Party,
   State_2 = df2$State,
   Year_2 = df2$Start_Year,
+  Party_2 = df2$Party,
   stringsAsFactors = FALSE
@@
 
-# Create the table
-kable(compact_df, col.names = c("State", "Last Time Voted for a Different Party",
-                                "State", "Last Time Voted for a Different Party"))
+# Function to assign color names based on party
+assign_color <- function(party) {
+  ifelse(party == "Rep", "#DD1717", "#0F4392")
+}
+
+# Apply cell_spec to each cell that needs coloring
+compact_df <- compact_df %>%
+  mutate(
+    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
+    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
+    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE),
+    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE)
+  )
+
+# Create and style the table
+styled_table <- compact_df %>%
+  select(State_1, Year_1, State_2, Year_2) %>%
+  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c",
+        col.names = c("State", "Last Time Voted for a Different Party",
+                      "State", "Last Time Voted for a Different Party")) %>%
+  kable_styling(latex_options = c("striped"), full_width = FALSE, position = "left") %>%
+  row_spec(0, bold = TRUE) %>%
+  footnote(general = " Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.")
+
+# Display the table
+styled_table
+
 ```
